### 2018-06-25 projects/dullemond/paper_dusttraps/ms.tex
New version of paper on dust traps by Kees et al.

%\documentclass[referee]{aa}
\documentclass{aa}
%\documentclass{mn2e}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{natbib}
\usepackage{xspace}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{hyperref}
\def\remark#1{{{\bf remark:} \bf #1}}
\def\action#1{{\bf #1}}
\def\putin#1{{\it #1}}
\def\revised#1{{\bf #1}}
\newcommand{\todo}[1]{\textcolor{red}{\bf #1}}
%
\def\aap{A\&A}
\def\apss{A\&SS}
\def\pasp{Proc.Astr.Soc.Pacific}
\def\pasj{Publ.Astr.Soc.Japan}
\def\araa{AnnRevA\& A}
\def\mnras{MNRAS}
\def\icarus{Icarus}
\def\nat{Nature}
\def\apj{ApJ}
\def\aj{AJ}
\def\apjl{ApJL}
\def\apjs{ApJS}
\def\jqsrt{JQSRT}
\def\max{\mathrm{max}}
\def\erg{\hbox{erg}}
\def\sec{\hbox{s}}
\def\Hz{\hbox{Hz}}
\def\gram{\hbox{g}}
\def\cm{\hbox{cm}}
\def\km{\hbox{km}}
\def\kel{\hbox{K}}
\def\KeV{\hbox{KeV}}
\def\MeV{\hbox{MeV}}
\def\eV{\hbox{eV}}
\def\yr{\hbox{yr}}
\def\AU{\hbox{AU}}
\def\ster{\hbox{ster}}
%\def\in{\mathrm{in}}
\def\out{\mathrm{out}}
\def\eff{\mathrm{eff}}
\def\rad{\mathrm{rad}}
\def\vrt{\mathrm{vert}}
\def\vset{\ensuremath{v_\mathrm{sett}}\xspace}
\def\fact{\ensuremath{\xi}\xspace}
\def\cs{\ensuremath{c_\mathrm{s}}\xspace}
\def\siggas{\ensuremath{\sigma_\mathrm{g}}\xspace}
\def\sigcoag{\ensuremath{\sigma_\mathrm{c}}\xspace}
\def\Tgas{\ensuremath{T_\mathrm{gas}}\xspace}
%\def\rhogas{\ensuremath{\rho_\mathrm{gas}}\xspace}
\def\rhogas{\ensuremath{\rho}\xspace}
\def\rhodust{\ensuremath{\rho_\mathrm{d}}\xspace}
\def\mugas{\ensuremath{\mu_\mathrm{gas}}\xspace}
\def\mp{\ensuremath{m_\mathrm{p}}\xspace}
\def\Hp{\ensuremath{H_\mathrm{p}}\xspace}
\def\Hs{\ensuremath{H_\mathrm{s}}\xspace}
\def\Omegak{\ensuremath{\Omega_\mathrm{K}}\xspace}
\def\Mstar{\ensuremath{M_\star}\xspace}
\def\Sc{\ensuremath{\mathrm{Sc}}\xspace}
\def\St{\ensuremath{\mathrm{St}}\xspace}
\def\Pt{\ensuremath{\mathrm{Pt}}\xspace}
\def\comma{\,,}
\def\fullstop{\,.}
%
\def\submitted{submitted\xspace}
\def\inprep{in prep.\xspace}
%
\begin{document}
%\thesaurus{02.01.2,08.03.4,08.06.2,08.16.5,13.09.6}
\title{Dust trapping in thin-ringed protoplanetary disks}  \titlerunning{Thin-ringed disks}
\authorrunning{Dullemond, Birnstiel, Troncoso, Perez, Andrews, Isella, Huang, Zhu, Guzman, Benisty, Carpenter}
\author{C.P.~Dullemond$^{1}$, T.~Birnstiel$^{2}$, N.~Troncoso$^{3}$, L.~Perez$^{3}$, S.~Andrews$^{4}$, A.~Isella$^{5}$, J.~Huang$^{4}$, Z.~Zhu$^{6}$, V.~Guzm\'an$^{7}$, M.~Benisty$^{8}$, J.~Carpenter$^{9}$}
\institute{
  (1) Zentrum f\"ur Astronomie, Heidelberg University, Albert Ueberle Str.~2, 69120 Heidelberg, Germany\\
  (2) {\bf XXXX}
  (3) {\bf XXXX}
  (4) {\bf XXXX}
  (5) {\bf XXXX}
  (6) {\bf XXXX}
} \date{\today}

\abstract{A large fraction of the protoplanetary disks observed with the ALMA
  Large Programme {\bf XXXX} display multiple well-defined and nearly perfectly
  circular rings in the continuum, in many cases with substantial peak-to-valley
  contrast. Several of these rings are very narrow in radial extent.
  In this paper we analyze these results using the assumption that
  these dust rings are caused by dust trapping in radial pressure bumps. We
  model this process in a 1-D axisymmetric way, initially with a simple analytic
  model, then with a more detailed numerical model. We find that .....
%  including time-dependent viscous disk evolution coupled to time-dependent
%  radial dust drift. We perform model calculations without and with dust growth,
%  the latter with a simplified dust growth recipe. We test several scenarios
%  for the origin of the pressure bumps, including planetary gaps and dead zones.
{\bf CONCLUSIONS}
}

\maketitle

\begin{keywords}
accretion, accretion disks -- circumstellar matter 
-- stars: formation, pre-main-sequence -- infrared: stars 
\end{keywords}

\section{Introduction}
The concept of dust trapping in local pressure maxima has become a central theme
in studies of planet formation and protoplanetary disk evolution, because it might
provide an elegant solution to several problems in these fields of
study. Theories of planet formation are plagued by the ``radial drift barrier'':
the problem that, as dust aggregates grow by coagulation, they tend to radially
drift toward the star before they reach planetsimal size
\citep[e.g.][]{2010A&A...513A..79B}. A natural solution to this problem could be
the trapping of dust particles in local pressure maxima
\citep{1972fpp..conf..211W, 2007ApJ...664L..55K, 1995A&A...295L...1B,
  1997Icar..128..213K}. Not only does this process prevent excessive radial
drift of dust particles, it also tends to concentrate the dust into small
volumes and high dust-to-gas ratios, which is beneficial to planet formation.
From an observational perspective, the radial drift
problem manifests itself by the presence of large grains in the outer regions of
protoplanetary disks \citep{2003A&A...403..323T, 2009ApJ...700.1502A}, which
appears to be in conflict with theoretical predictions
\citep{2007A&A...469.1169B}. One possible solution to this observational
conundrum could be that the disks are much more massive in the gas than
previously suspected, leading to a higher gas friction for millimeter grains and
thus longer drift time scales \citep{2017ApJ...840...93P}. Another explanation is
to invoke dust traps. The most striking observational evidence for
dust trapping seems to come from large transitional disks, which feature giant
dust rings, sometimes lopsided, in which large quantities of dust appears to be
concentrated \citep{2013Natur.493..191C, 2013Sci...340.1199V}. These observations
appear to be well explained by the dust trapping scenario
\citet{2012A&A...545A..81P}. But these transitional disks seem to be rather
violent environments, possibly with strong warps \citep{2015ApJ...798L..44M,
  2017A&A...597A..42B} and companion-induced spirals
\citep{2016ApJ...816L..12D}. For more ``normal'' protoplanetary disk the dust
traps would have to be more subtle. \cite{2012A&A...538A.114P} explored the
possibility that the disk contains many axisymmetric local pressure maxima, and
calculated how the dust drift and growth would behave under such conditions. It
was found that, if the pressure bumps are strong enough, the dust trapping can
keep a sufficient fraction of the dust mass at large distances from the star to
explain the observed dust millimeter flux. It would leave, however, a detectable
pattern of rings that should be discernable with ALMA observations.  Since the
multi-ringed disk observation of HL Tau \citep{2015ApJ...808L...3A} a number of
such multi-ringed disks have been detected \citep{2016ApJ...820L..40A,
  2016PhRvL.117y1101I, 2017ApJ...851L..23C, 2017A&A...600A..72F,
  2018A&A...610A..24F}. It is therefore very tempting to see also these
multi-ringed disks as evidence for dust trapping, and as an explanation for
the retention of dust in the outer regions of protoplanetary disks.

The data from the ALMA Large Programme 2016.1.00484.L offers an exciting new
opportunity to put this concept to the test, and to put constraints on the
physics of dust trapping in axisymmetric pressure maxima. This is an opportunity
which we explore in this paper.

If we assume that the rings seen in our data are caused by dust trapping, the
question arises: what constraints do the data impose on the physics of dust
trapping? To explore this we start by studying the rings individually, assuming
that the dust does not escape from the ring. This makes it possible to look for
a steady-state dust trapping solution in which the radial drift forces (that push
the dust to the pressure peak) are balanced by turbulent mixing (that tends to
smear out the dust away from the pressure peak). In Section
\ref{sec-steady-state-analytic-trap-model} we will construct a very simplified
analytic dust trapping model, and confront this with the most well-isolated
rings from our sample. {\bf Describe the next sections.} But before we do the
modeling, we review, in Section \ref{sec-data}, the key aspects of the data from
our ALMA Large Programme that we compare to our models.

\section{The high-contrast rings of AS 209, Elias 24, HD 163296, GW Lup and HD 143006}
\label{sec-data}
%
In this paper we focus on those sources of the ALMA Large Programme
that show high-contrast and radially thin rings that are separated by deep
valleys, and that are sufficiently face-on to not have to worry about 3-D
line-of-sight issues. These are AS 209, Elias 24, HD 163296, GW Lup and
HD 143006. Their stellar parameters are given in Table \ref{tab-stellar-params}.

\begin{table}
\begin{center}
\begin{tabular}{|c|ccc|}
\hline
\hline
Source     & $d\,\mathrm{[pc]}$ & $M_{*}/M_{\odot}$ & $L_{*}/L_{\odot}$ \\
\hline
AS 209     & 121              & 0.912        &  0.631 \\
Elias 24   & 136              & 0.776        &  6.03  \\
HD 163296  & 101              & 2.14         &  25.7  \\
GW Lup     & 155              & 0.457        &  0.331 \\
HD 143006  & 166              & 0.813        &  0.871 \\
\hline
\hline
\end{tabular}
\end{center}
\caption{\label{tab-stellar-params}The stellar parameters for the
  stars studied in this paper. Distance is in parsec and mass and luminosity
  are in units of the solar values.}
\end{table}

A gallery of these sources is shown in Fig.~\ref{fig-obs-images}.
For an overview of the ALMA Large Programme we refer to {\bf Andrews et
  al.~XXXX}, and for an in-depth discussion on the data of the individual
sources we refer to {\bf [LIST HERE THE PAPERS]}.

\begin{figure*}
\centerline{\includegraphics[width=0.95\textwidth]{gallery_rings.pdf}}
\caption{\label{fig-obs-images}The continuum maps in
  band 6 of the five disks in our sample which have the most pronounced rings.
  The eight highest contrast rings, which are the topic of this paper,
  are marked in the images. For a detailed description of these
  data, see {\bf Guzm\'an et al.~(2018)} for AS 209, {\bf XXXX}}
\end{figure*}

The high-contrast rings of these sources provide ``clean laboratories''
for testing the theory of dust trapping, both in a ring-by-ring manner, as well
as by comparing to global dust drift models. Fig.\ \ref{fig-obs-profiles}
shows the radial profile (deprojected for
inclination) of the thermal emission of the dust of the five disks. The
procedure used to extract these radial profiles from the continuum maps
is described in {\bf [Paper Jane Huang, but note that for HD 163296 and
HD 143006 we removed the blob]}. The
brightness is expressed as linear brightness temperature
$T_{\mathrm{bright,linear}}=(c^2/2k_B)I_\nu/\nu^2$, where $I_\nu$ is the
intensity, $c$ the light speed, $k_B$ the Boltzmann constant, and $\nu$ the
frequency\footnote{It should be noted that the full (non-linear) brightness temperature,
defined such that $B_\nu(T_{\mathrm{bright,full}})=I_\nu$, with $B_\nu(T)$ the
Planck function, can have very different values, in particular at low
brightness. As a result $T_{\mathrm{bright,full}}(r)$ tends to vastly
underestimate the contrast between the rings and the inter-ring
regions. Therefore, in this paper we plot the brightness using the linear
brightness temperature, which is proportional to the intensity in Jy/beam.}.

\begin{figure*}
\centerline{\includegraphics[width=0.85\textwidth]{gallery_profiles.pdf}}
\caption{\label{fig-obs-profiles}The linear brightness temperature in
  band 6 of the five disks in our sample which have the most pronounced rings.
  The vertical axis is logarithmic to better show the contrast.
  The eight highest contrast rings are fitted by a Gaussian profile, shown
  as the solid inverse parabolas. The dotted inverse parabolas are Gaussians
  with the width of the ALMA beam. For a detailed description of these
  data, see {\bf Guzm\'an et al.~(2018)} for AS 209, {\bf XXXX}}
\end{figure*}

The eight rings which we study in this paper (see annotations in
Fig.~\ref{fig-obs-images}) are fitted with a Gaussian profile, shown as inverse
parabolas in Fig.\ \ref{fig-obs-profiles}. The fitting procedure was done in the
image plane, in order to make it possible to fit to a limited radial range,
i.e.\ selecting a single ring. Details of this fitting procedure, and tests
using the ALMA simulator, are described in Appendix {\bf XXXXX}. In
Fig.~\ref{fig-obs-gaussfits} these gaussian fits are shown in close-up. The
parameters of the best-fitting Gaussians are listed in Table
\ref{tab-gauss-params}.

\begin{figure*}
\centerline{\includegraphics[width=0.85\textwidth]{gaussfits_allrings.pdf}}
\caption{\label{fig-obs-gaussfits}Gaussian fits to the eight rings of this
  paper. The blue curves are the observations, the green curves are the best fit
  Gaussian profiles. The ``fit range'' bar shows the radial range within which
  the Gauss was fitted to the data. The fit range was chosen to fit the part of
  the curve that, by eye, most resembles a Gaussian. The ``beam'' bar shows the
  FWHM beam size of the observations. The grey band around the blue curve shows
  the estimated uncertainty of the data.}
\end{figure*}

\begin{figure*}
  \centerline{
    %\includegraphics[width=0.3\textwidth]{AS_209_doublegauss.pdf}
    \includegraphics[width=0.42\textwidth]{HD_163296_doublegauss.pdf}
    \includegraphics[width=0.42\textwidth]{HD_143006_doublegauss.pdf}
  }
  \caption{\label{fig-obs-doublegauss}The sum of the two Gaussian fits
    for the two sources with two partly overlapping rings: HD 163296 and
    HD 143006.}
\end{figure*}

\begin{table*}
\begin{center}
\begin{tabular}{|cc|cc|ccccc|ccccc|}
\hline
\hline
%Source & Ring & $r_i\, [\mathrm{au}]$ & $r_o\, [\mathrm{au}]$ & $A\, [\mathrm{K}]$ & $r\, [\mathrm{au}]$ & $\mathrm{fwhm}\, [\mathrm{mas}]$ & $\sigma\, [\mathrm{au}]$ & $w_d\, [\mathrm{au}]$ & $T_{\mathrm{model}}\,\mathrm{[K]}$ & $w_d/h_p$ & $M_{\mathrm{d,ri}} [M_{\oplus}]$ & $M_{\mathrm{d,iv}} [M_{\oplus}]$\\
Source & Ring & $r_{\mathrm{in}}$ & $r_{\mathrm{out}}$ & $A$ & $r_0$ & $\mathrm{fwhm}$ & $\sigma$ & $w_d$ & $T_{\mathrm{model}}$ & $T_{\mathrm{model,br,lin}}$ & $w_d/h_p$ & $M_{\mathrm{d,ri}}$ & $M_{\mathrm{d,iv}}$\\
            & & $[\mathrm{au}]$ & $[\mathrm{au}]$ & $[\mathrm{K}]$ & $[\mathrm{au}]$ & $[\mathrm{mas}]$ & $[\mathrm{au}]$ & $[\mathrm{au}]$ & $\mathrm{[K]}$ & $\mathrm{[K]}$ &  & $[M_{\oplus}]$ & $[M_{\oplus}]$\\
\hline
% NOTE: This table is made automatically with python_make_tables/make_latex_table.py
AS 209     & 1 &  69 &  79 &   3.64 &  74.13 &  75.42 &   3.88 &   3.43 &  12.9 &   8.2 &   0.7 &  44.4 &   2.41\\
AS 209     & 2 & 115 & 125 &   2.95 & 120.31 &  90.28 &   4.64 &   4.28 &  10.1 &   5.6 &   0.5 & 101.9 &   5.54\\
Elias 24   & 1 &  72 &  82 &   5.22 &  76.85 &  87.99 &   5.08 &   4.66 &  22.3 &  17.2 &   0.6 &  41.1 &   2.23\\
HD 163296  & 1 &  60 &  76 &   8.48 &  67.41 & 171.73 &   7.36 &   7.21 &  34.2 &  29.0 &   1.6 &  50.4 &   2.74\\
HD 163296  & 2 &  94 & 104 &   4.92 &  99.86 & 133.56 &   5.73 &   5.53 &  28.1 &  22.9 &   0.8 &  42.5 &   2.31\\
GW Lup     & 1 &  79 &  89 &   1.48 &  85.13 &  72.08 &   4.74 &   4.15 &  10.3 &   5.7 &   0.6 &  36.3 &   1.97\\
HD 143006  & 1 &  35 &  45 &   3.21 &  41.05 &  71.76 &   5.06 &   4.42 &  18.8 &  13.8 &   1.7 &  16.7 &   0.91\\
HD 143006  & 2 &  59 &  72 &   2.48 &  64.83 & 111.83 &   7.88 &   7.49 &  15.0 &  10.1 &   1.7 &  43.5 &   2.36\\
\hline
\hline
\end{tabular}
\end{center}
\caption{\label{tab-gauss-params}The model parameters for the Gaussian ring fits
  in Figs.~\ref{fig-obs-profiles} and \ref{fig-obs-gaussfits}. The values $r_{\mathrm{in}}$
  and $r_{\mathrm{out}}$ are the inner and outer boundary of the fitting domain. $A$ is the
  peak brightness temperature of the ring, $r_0$ is its radius in $\mathrm{au}$,
  $\mathrm{fwhm}$ is the width expressed as full-width-at-half-maximum in units
  of milliarcsecond, $\sigma$ is its width expressed as standard deviation in
  units of $\mathrm{au}$ and $w_d$ is the width of the underlying (deconvolved)
  dust emission profile, also expressed as standard deviation in units of
  $\mathrm{au}$. The last five columns are derived from a simple disk model
  assumption: $T_{\mathrm{model}}$ is the midplane temperature computed from
  Eq.~(\ref{eq-disk-temperature-model}) assuming a flaring angle of
  $\varphi=0.02$. From that one can derive the corresponding linear brightness
  temerature $T_{\mathrm{model,br,lin}}=B_\nu(T_{\mathrm{model}})c^2/2\nu^2$ with
  $\nu=c/0.13\,\mathrm{cm}$ is the frequency belonging to the wavelength of
  0.13 $\mathrm{cm}$ of ALMA band 6.
  The third-to-last column expresses the deconvolved dust ring
  width in units of the disk pressure scale height $h_p$ computed from
  $T_{\mathrm{model}}$. The last two columns are the computed dust mass in the
  ring in units of Earth masses, assuming the small grain absorption opacity
  $\kappa_{\mathrm{abs,ri}}(\lambda=0.13\,\mathrm{cm})=1.70\,\mathrm{cm}^2/\mathrm{g}$
  of \citet{2010A&A...512A..15R} ($M_{\mathrm{d,ri}}$) and the corresponding
  value
  $\kappa_{\mathrm{abs,ri}}(\lambda=0.13\,\mathrm{cm})=31.3\,\mathrm{cm}^2/\mathrm{g}$
  of \citet{1997MNRAS.291..121I} ($M_{\mathrm{d,iv}}$). The stellar parameter
  are taken from Table \ref{tab-stellar-params}.}
\end{table*}

The observed rings are the result of the thermal emission of a dust ring convolved
with the ALMA beam. The standard deviation width $\sigma$ of the fitted Gaussian
profile is given in Table \ref{tab-gauss-params}.  To obtain the width of the
underlying dust ring we have to deconvolve. Assuming a Gaussian beam and a
Gaussian dust ring, we can use the rule of the convolution of two Gaussians, and
obtain the width $w_d$ of the dust ring
\begin{equation}
w_d=\sqrt{\sigma^2-b^2}
\end{equation}
where $b$ is the beam width expressed as standard deviation in units of
$\mathrm{au}$ ($b=d_{\mathrm{pc}}b_{\mathrm{fwhm,as}}/2.355$, where
$b_{\mathrm{fwhm,as}}$ is the beam full width at half maximum in units of
arcseconds, and $d_{\mathrm{pc}}$ is the distance to the source in units of
parsec). The width of the dust rings
in units of the local pressure scale height $h_p$ (last column of Table
\ref{tab-gauss-params}) can only be roughly estimated, because we do not know
the disk midplane temperature. Our estimate is made as follows:
\begin{equation}\label{eq-disk-temperature-model}
T_{\mathrm{model}}(r) = \left(\frac{\tfrac{1}{2}\varphi L_{*}}{4\pi r^2\sigma_{\mathrm{SB}}}\right)^{1/4}
\end{equation}
where $\sigma_{\mathrm{SB}}$ is the Stefan-Boltzmann constant and $\varphi$ is
the so-called flaring angle \citep[e.g.][]{1997ApJ...490..368C,
  1998ApJ...500..411D, 2001ApJ...560..957D}. The pressure scale height then
follows from $h_p=\sqrt{k_BT_{\mathrm{model}}r^3/\mu m_pGM_{*}}$ with $k_B$ the
Boltzmann constant, $m_p$ the proton mass, $G$ the gravitational constant and
$mu=2.3$ the mean molecular weight in atomic units. The stellar parameters are
taken from Table \ref{tab-stellar-params}. We take the flaring angle to be
$\varphi=0.02$ which is an estimate based on typical values from models.  But
this value can easily be wrong by a factor of four or so. The fact that the
pressure scale height goes as $T_{\mathrm{model}}^{1/2}\propto \varphi^{1/8}$
means, however, that such an error only affects $h_p$ mildly.

We see from Table \ref{tab-gauss-params} that some rings are narrower than the
(estimated) pressure scale height $h_p$, while others are broader. This
comparison is important because long-lived pressure bumps in the gas cannot be
radially narrower than about one pressure scale height. If a thermal emission
ring produced by the dust is substantially narrower than $h_p$, then some kind
of dust trapping must have taken place. If this were not the case, then the dust
emission profile would have the same shape as the gas midplane density, and thus
the gas midplane pressure bump would then be a narrower than $h_p$.

We see that 5 out of 8 rings are narrower than $h_p$, but not by much. None are
less than half the width of the pressure scale height of the gas. This is
perhaps slightly disappointing, because it means that these results are no
irrefutable proof that dust trapping is playing a role in the shaping of these
rings. But it is strong evidence for that nonetheless.

For these observations the fwhm beam size is of the order of 35 milliarcseconds
{\bf [CHECK AND MAKE MORE ACCURATE]}. This number can be compared to the fwhm
ring width in Table \ref{tab-gauss-params}, and it shows that the widths of all
the rings are spatially resolved. Some are only marginally resolved (AS 209
rings 1 and 2, GW Lup and the ring 1 of HD 143006), but the data are
inconsistent with radially unresolved rings. This is confirmed by more detailed
analysis in the uv-plane ({\bf REFS TO Guzm\'an et al.~(2018)}).

These Gaussian profiles fit the observed data remarkably well, as can be
best seen in Fig.~\ref{fig-obs-gaussfits}. 
For the broad rings of HD 163296 the closeness to a Gaussian profile is
particularly striking, given that these rings are resolved by many beam
widths. In particular in this source there is therefore no doubt that the
Gaussian profile is a reflection of a real phenomenon, and is not caused by
unknown observational or data reduction issues.  But also the rings of AS 209,
while not as well-resolved as those of HD 163296, are remarkably close to
Gaussian.

For Elias 24 the ring fits a Gaussian near the peak, but deviates from Gaussian
already less half a width from the center. Likewise for the rings of HD
143006. The ring around GW Lup is an intermediate case, which fits a Gauss well
on the inner flank, but rather badly on the outer flank.

In all ring profiles shown in Fig.~\ref{fig-obs-gaussfits} the observed profile
rises above the Gaussian fit at some point in the wings. For the inner flanks of
ring 1 of Elias 24 and ring 1 of HD 143006, as well as for the outer flanks of
ring 2 of AS 209, ring 2 of HD 163296, the ring of GW Lup and ring 2 of HD
143006 the excess above the Gaussian gradually increases away from the peak of
the Gaussian. Put differently, the profiles appear to deviate from Gaussian
shape to pyramid-shape (or Lorenzian shape) in the flanks. This cannot be simply
the addition of a constant background, because that would preserve the Gaussian
shape, and merely lift it higher. Instead, while the core of the profile can clearly
be well-described by a Gaussian, in the flanks there is a clear transition
to a different, wider, shape.

A similar effect occurs between the two rings in HD 163296 and HD 143006, where
in particular for HD 143006 it is clear from Fig.~\ref{fig-obs-doublegauss} that
it cannot be just the effect of the sum of the two Gaussians.

It is not immediately clear whether this is relevant or just chance. It lies in
the nature of fitting a Gaussian profile to something non-Gaussian that there
will be a region close to the peak where the curve fits the Gaussian reasonably
well, while the deviation will increase the farther away from the peak one
looks. This danger is particularly large in the present case, since the fitting
range was chosen to maximize the similarity to the Gauss near the
peak. Nevertheless, it clearly signals a deviation from Gaussian.

Given the closeness of the profiles to a Gaussian, we can use the Gaussian fit
to make estimates of the total mass of dust trapped in each of these rings. The
procedure is described in Appendix \ref{sec-computing-dust-mass}, and the
results, for the two opacity choices we make, are listed in Table
\ref{tab-gauss-params}. In making these estimates we have assumed the dust to be
made up of micron-sized dust, which is likely too small. The mass based on the
Ivezic opacity, $M_{\mathrm{d,iv}}$, will increase for $a>\lambda/2\pi$ by a
factor of $2\pi\,a/\lambda$ with $\lambda=0.13\,\mathrm{cm}$. For the mass based
on the Ricci opacity the factor first decreases when $a$ approaches the opacity
peak around $a=\lambda/2\pi$ and then increases for $a>\lambda/2\pi$. Both
behaviors can be seen as the inverse behavior of the opacities shown in
Fig.~\ref{fig-opacities}. For ever increasing grain size estimate, the mass
estimate will increase indefinitely for both opacity choices.

\section{The rings as dust traps}
\label{sec-rings-as-dust-traps}
%
Given the well separated narrow rings in the above sources, we can analyze these
rings on a one-by-one basis. The hypothesis we are going to test is that the
rings are caused by dust trapping in axisymmetric pressure bumps. For simplicity
we will assume that the radial gas pressure profile is fixed in time, and there
is no back reaction of the dust onto the gas. The pressure bump is assumed to be
so strong that the dust trapping in these rings is nearly perfect: no dust
escapes. We then expect that the dust distribution finds an equilibrium between
dust drift and turbulent spreading. Low turbulence will lead to thinner dust
rings than high turbulence. For simplicity we shall assume the gas pressure bump
to have a Gaussian radial profile with its peak at radius $r_0$ and width $w\ll
r_0$. This problem can be solved analytically.

\subsection{Model}
\label{sec-model-dusttrap-rings}
%
Consider the following radial Gaussian profile for the
pressure at the disk midplane:
\begin{equation}\label{eq-gaussian-pressure-bump}
p(r) = p_0 \exp\left(-\frac{(r-r_0)^2}{2w^2}\right)
\end{equation}
We will discuss the hydrodynamic stability of a gas
distribution arranged in this way in Section \ref{sec-gas-ring-stability}. But
for now we will not be concerned with that and we will assume the
pressure-profile of Eq.~(\ref{eq-gaussian-pressure-bump}) to be given and fixed.

In Appendix \ref{sec-steady-state-analytic-trap-model} we derive that, for our
purposes, a good approximation of the stationary dust distribution within the
dust trap is:
\begin{equation}\label{eq-analytic-sol-radial-trapping-summary}
\Sigma_{\mathrm{d}}(r) = \Sigma_{\mathrm{d0}} \exp\left(-\frac{(r-r_0)^2}{2w_{\mathrm{d}}^2}\right)
\end{equation}
where
\begin{equation}\label{eq-wd-afo-w-psi}
  w_{\mathrm{d}} = w\,\left(1+\psi^{-2}\right)^{-1/2}
\end{equation}
with $\psi$ given by
\begin{equation}\label{eq-psi-afo-alpha-sc-st}
\psi = \sqrt{\frac{\alpha_{\mathrm{turb}}}{\mathrm{Sc}\,\mathrm{St}}}
\end{equation}
Here $\mathrm{St}$ is the Stokes number of the dust particles
(Eq.~\ref{eq-definition-stokes-number}), $\mathrm{Sc}$ is the Schmidt number of
the turbulence in the gas, and $\alpha_{\mathrm{turb}}$ is the usual turbulence
parameter. Note that this solution is for a single grain size.

For large grains and/or weak turbulence one finds $\psi\ll 1$, which leads to
$w_{\mathrm{d}}\ll w$. In this case the dust is strongly trapped near the peak
of the pressure bump. The opposite is the case for small grains and/or strong
turbulence, for which one gets $\psi\gg 1$, which leads to
$w_{\mathrm{d}}\rightarrow w$. In this case the trapping is very weak and the
dust-to-gas ratio within the pressure bump stays constant. Only in the very
wings of the Gaussian pressure bump is dust trapping significant, but this
effect can only be studied using the more accurate non-gaussian solution of
Eq.~(\ref{eq-analytic-sol-radial-trapping-better}). We will return to the dust
trapping effect in the wings in Section {\bf XXXXX}.

Eq.~(\ref{eq-analytic-sol-radial-trapping-summary}) has only two parameters:
$\Sigma_{\mathrm{d0}}$ and $w_{\mathrm{d}}$. Both can be fairly directly
measured from the observations, though $\Sigma_{\mathrm{d0}}$ will depend on the
opacity, and hence on the unknown grain size (see discussion in Appendix
\ref{sec-computing-dust-mass}). The values of $w_{\mathrm{d}}$ for
the rings in our sample can be directly taken from Table \ref{tab-gauss-params}.

The width of the dust ring $w_{\mathrm{d}}$ is physically set by
$\alpha_{\mathrm{turb}}$, $\mathrm{Sc}$, $\mathrm{St}$ and $w$ through the above
equations. We therefore
have one observational value for four unknown parameters. This is heavily
degenerate. All we can do is to test if the measured value of
$w_{\mathrm{d}}$ is consistent with expected values of $\alpha_{\mathrm{turb}}$,
$\mathrm{Sc}$, $\mathrm{St}$ and $w$.

\subsection{Ranges of $\alpha_{\mathrm{turb}}$, $\mathrm{Sc}$, $\mathrm{St}$ and $w$}
\label{sec-ranges-of-params}
%
First of all, the Schmidt number $\mathrm{Sc}$ is merely a way to relate the
turbulent viscosity with the turbulent mixing. If we do not strive to learn
about the turbulent viscosity and instead are satisfied with learning only about
the turbulent mixing, then we are only interested in the combination
$\alpha_{\mathrm{turb}}/\mathrm{Sc}$. For simplicity we set $\mathrm{Sc}=1$.

The value of the turbulence parameter $\alpha_{\mathrm{turb}}$ is usually
considered to be between $10^{-6}\lesssim \alpha_{\mathrm{turb}}\lesssim
10^{-2}$.

The width of the pressure bump cannot be smaller than about a pressure
scale height, but also not smaller than the width of the dust ring.
Therefore $w_{\mathrm{min}}=\mathrm{max}(h_p,w_{\mathrm{d}})$.
In the case of the double rings (AS 209, HD 163296 and HD 143006),
the full-width-at-half-maximum $2.355\,w$ should not exceed the radial
separation of the rings. For the two single ring sources we take the deepest
point of the gap to the inside of the ring as the upper limit on the
half-width-at-half-maximum $1.178\,w$. These lower and upper limits on $w$
are listed in Table \ref{tab-ring-model-limits}.

The Stokes number $\mathrm{St}$ can be any value. But it is
directly related to the grain size $a$ and the gas density
$\rho_{\mathrm{g}}$. If we have observational constraints on $a$ and a good
estimate of the gas density $\rho_{\mathrm{g}}$, then we can eliminate this
uncertainty, and we are left with two unknown parameters
($\alpha_{\mathrm{turb}}$ and $w$) for one measurement ($w_{\mathrm{d}}$).
Unfortunately, while estimating $a$ from observations may be doable, it is far
more difficult to estimate $\rho_{\mathrm{g}}$. Standard disk gas mass estimates
are of limited use, as they are based on measuring the dust mass and multiplying
it by the estimated gas-to-dust ratio. Since we are testing the hypothesis of
dust trapping, we cannot assume a standard gas-to-dust ratio.

One can, however, set an upper bound on $\rho_{\mathrm{g}}$ by demanding that
the disk is gravitationally stable, i.e.\ that the Toomre parameter obeys
\begin{equation}
Q_{\mathrm{Toomre}}\equiv \frac{c_s\Omega_K}{\pi G \Sigma_{\mathrm{g}}} >2
\end{equation}
Here $c_s=\sqrt{k_BT/\mu m_p}$ is the isothermal sound speed, with $k_B$ the
Boltzmann constant, $m_p$ the proton mass, and $mu=2.3$ the mean molecular
weight in atomic units. $\Omega_K=\sqrt{GM_{*}/r^3}$ is the Kepler frequency,
$G$ is the gravitational constant and, finally, $\Sigma_{\mathrm{g}}$ the gas
surface density. The gas surface density and midplane density are related via
the pressure scale height $h_p=c_s/\Omega_K$ via the identity
$\Sigma_{\mathrm{g}}=\sqrt{2\pi}h_p\rho_{\mathrm{g}}$. Taking the disk midplane
temperature from Table \ref{tab-gauss-params}, which was calculated using
Eq.~(\ref{eq-disk-temperature-model}), we can compute the upper limits on
$\rho_{\mathrm{g}}$ for all of the rings. The results are given in Table
\ref{tab-ring-model-limits}.

\begin{table*}
\begin{center}
\begin{tabular}{|cc|cccccc|cc|}
\hline
\hline
Source     & Ring & $w_{\mathrm{min}}$ & $w_{\mathrm{max}}$ & $\rho_{\mathrm{g,min,iv}}$ & $\rho_{\mathrm{g,min,ri}}$ & $\rho_{\mathrm{g,max}}$ & $a_{\mathrm{max}}$ & $(\alpha/\mathrm{St})_{\mathrm{min}}$ & $(\alpha/\mathrm{St})_{\mathrm{max}}$\\
           &      & $[\mathrm{au}]$  & $[\mathrm{au}]$  & $[\mathrm{g}/\mathrm{cm}^3]$ & $[\mathrm{g}/\mathrm{cm}^3]$ & $[\mathrm{g}/\mathrm{cm}^3]$ & $[\mathrm{cm}]$ & $\equiv\psi^{-2}_{\mathrm{max}}$ & $\equiv\psi^{-2}_{\mathrm{min}}$\\
\hline
AS 209     & 1 &  4.83 & 19.6 & 7.9e-17 & 1.4e-15 & 8.4e-14 & 11.1 & 3.2e-02 & 1.0e+00 \\
AS 209     & 2 &  8.85 & 19.6 & 5.1e-17 & 9.3e-16 & 2.0e-14 &  4.0 & 5.0e-02 & 3.0e-01 \\
Elias 24   & 1 &  7.26 & 17.1 & 3.6e-17 & 6.5e-16 & 6.4e-14 & 18.8 & 8.0e-02 & 7.0e-01 \\
HD 163296  & 1 &  7.21$^{*}$ & 13.8 & 5.6e-17 & 1.0e-15 & 2.6e-13 & 48.7 & 3.8e-01 & -- \\
HD 163296  & 2 &  7.27 & 13.8 & 2.5e-17 & 4.6e-16 & 8.1e-14 & 33.4 & 1.9e-01 & 1.4e+00 \\
GW Lup     & 1 &  7.49 &  9.9 & 2.9e-17 & 5.4e-16 & 2.8e-14 &  9.8 & 2.1e-01 & 4.4e-01 \\
HD 143006  & 1 &  4.42$^{*}$ & 10.1 & 7.8e-17 & 1.4e-15 & 4.4e-13 & 58.9 & 2.4e-01 & -- \\
HD 143006  & 2 &  7.49$^{*}$ & 10.1 & 4.6e-17 & 8.5e-16 & 1.1e-13 & 25.1 & 1.2e+00 & -- \\
\hline
\hline
\end{tabular}
\end{center}
\caption{\label{tab-ring-model-limits}Limits on the free parameters of the dust
  trapping model. The lower limit to the pressure bump width $w_{\mathrm{min}}$
  is the pressure scale height $h_p$. If, however, the width of the dust ring
  $w_d>h_p$, then the lower limit is $w_d$ (marked with the symbol $^{*}$).  The
  upper limit $w_{\mathrm{max}}$ is derived from the separation between the
  rings (for AS 209, HD 163296 and HD 143006) or from the separation of the ring
  to the nearest minimum (for Elias 24 and GW Lup). The two lower limits on the
  gas density $\rho_{\mathrm{g}}$ were derived by demanding
  $\Sigma_{\mathrm{g}}\gtrsim \Sigma_{\mathrm{d}}$, where $\Sigma_{\mathrm{d}}$
  was computed in Appendix \ref{sec-computing-dust-mass} using two different
  opacities: that of Ivezic et al. (leading to $\rho_{\mathrm{g,min,iv}}$) and
  that of Ricci et al.\ (leading to $\rho_{\mathrm{g,min,ri}}$). The column
  marked with $\rho_{\mathrm{g,max}}$ is the upper limit on the midplane gas
  density derived from demanding that the gas disk is gravitationally
  stable. The $a_{\mathrm{max}}$ column is the maximum grain size for which the
  derived dust surface density together with the gas surface density remain
  gravitationally stable. The column $(\alpha/\mathrm{St})_{\mathrm{min}}\equiv
  \psi^{-2}_{\mathrm{max}}$ is a result of the analysis: the lower limit to
  $\alpha_{\mathrm{turb}}/\mathrm{St}$ derived from the ratio of
  $w_{\mathrm{max}}/w_{\mathrm{d}}$.}
\end{table*}

One can estimate a lower limit to the gas density by demanding that the gas
surface density must be at least as large as the dust surface density, since
dust trapping is unlikely to achieve a larger concentration of dust than
that. Depending on which of the two opacities we use (see Appendix
\ref{sec-computing-dust-mass}) we arrive at different estimates of
$\Sigma_{\mathrm{d}}(r=r_0)$. By demanding $\Sigma_{\mathrm{g}}\gtrsim
\Sigma_{\mathrm{d}}$, and using the pressure scale height $h_p$
and the formula $\rho_{\mathrm{g,midplane}}=\Sigma_{\mathrm{g}}/\sqrt{2\pi}h_p$,
we arrive at two lower limits of the midplane gas density, given in
Table \ref{tab-ring-model-limits}. It is likely that even for much
larger values of the midplane gas density the dust-gas mixture becomes
unstable to the streaming instability and other types of instabilities.
If we can convince ourselves that such instabilities are not operating
in these rings, then this puts a substantially stronger (i.e.~larger)
lower limit on the midplane gas density.

Also if the grains are much larger than $\lambda/(2\pi)\simeq 0.02\,\mathrm{cm}$,
the opacity drops and the resulting dust surface density estimate increases,
yielding larger lower limits to the gas density.

Along this line of thinking one can compute the largest grain radius for which
the Toomre parameter of the disk stays above 2. We use the Ivezic
opacity for that, which, for large grains, has similar values as the Ricci
opacity. This leads to values of several centimeters to half a meter (Table
\ref{tab-ring-model-limits}). Again these are conservative limits, with real
values likely to be substantially smaller.

\subsection{Application to the observed rings}
\label{sec-application-to-rings}
%
We now wish to apply the model of Subsection \ref{sec-model-dusttrap-rings}
with the limits on the parameter ranges derived in Subsection
\ref{sec-ranges-of-params} to the observed ring widths $w_d$ listed in
Table \ref{tab-gauss-params}.

Let us first choose the widest allowed pressure bump (i.e.\ taking
$w=w_{\mathrm{max}}$ from Table \ref{tab-ring-model-limits}). In that
case the dust rings are all narrower that the gas rings: $w_{\mathrm{d}}<w$,
which implies that the dust trapping is operational. We can derive the
values of $\psi$ directly from Eq.~(\ref{eq-wd-afo-w-psi}). And together
with Eq.~(\ref{eq-psi-afo-alpha-sc-st}) we can then directly derive the
following constraint on the unknown parameters $\alpha_{\mathrm{turb}}$
and $\mathrm{St}$, 
\begin{equation}
\frac{\mathrm{Sc}\,\mathrm{St}}{\alpha_{\mathrm{turb}}} \equiv \psi^2 = \left(\frac{w}{w_{\mathrm{d}}}\right)^2-1
\end{equation}
where, we recall, set $\mathrm{Sc}=1$. Given that this is the largest
possible value of $w$, and therefore the largest possible value
of $\psi$, this gives a lower limit to the ratio $\alpha_{\mathrm{turb}}/\mathrm{St}$:
\begin{equation}
\frac{\alpha_{\mathrm{turb}}}{\mathrm{St}} \gtrsim \frac{1}{\psi^2_{\mathrm{max}}} \qquad\hbox{($\psi_{\mathrm{max}}$ from largest $w$)}
\end{equation}
The resulting values are given in Table \ref{tab-ring-model-limits}.  These are
the first quantitative results of the application of the dust trapping model to
the data.

We can also derive an upper limit to the ratio
$\alpha_{\mathrm{turb}}/\mathrm{St}$, but only for the rings for which
$w_{\mathrm{d}}<h_p$. The reason is that for those rings with
$w_{\mathrm{d}}>h_p$ (marked with a $^{*}$ in Table \ref{tab-ring-model-limits})
the minimal pressure bump width is $w_{\mathrm{min}}=w_{\mathrm{d}}$, and the
dust ring is as wide as the pressure bump, implying that dust trapping is weak.
Any increase of $\alpha_{\mathrm{turb}}/\mathrm{St}$ will keep
$w_{\mathrm{d}}=w_{\mathrm{min}}$, so one cannot derive an upper limit on
$\alpha_{\mathrm{turb}}/\mathrm{St}$. For the rings with $w_{\mathrm{d}}<h_p$
(and thus $w_{\mathrm{min}}=h_p$) we can write
\begin{equation}
\frac{\alpha_{\mathrm{turb}}}{\mathrm{St}} \lesssim \frac{1}{\psi^2_{\mathrm{min}}} \qquad\hbox{($\psi_{\mathrm{min}}$ from smallest $w$)}
\end{equation}
The resulting values are given in Table \ref{tab-ring-model-limits}.  These are
the second quantitative results of the application of the dust trapping model to
the data.

If we will be able to measure the width of the gas pressure bump in an
independent way, for instance with molecular line observations, then instead of
the lower- and upper limits on $\alpha_{\mathrm{turb}}/\mathrm{St}$ we will be
able to directly measure $\alpha_{\mathrm{turb}}/\mathrm{St}$. However, reliable
measurements of the gas density are notoriously difficult. CO molecular lines
are usually emitted in the warm sub-surface layers of the disk, not near the
midplane (see e.g.~{\bf Weaver \& Isella companion paper as well as earlier
  literature}), and CO appears to be depleted compared to chemistry models
({\bf XXXX}). Other gas tracers are harder to measure and also not reliable
indicators of the gas density. Therefore we will, at least for now, have to
make do with the lower and upper limits on $\alpha_{\mathrm{turb}}/\mathrm{St}$
from Table \ref{tab-ring-model-limits}.

The next task is to convert from Stokes number $\mathrm{St}$ to grain size
$a$.

{\bf [TO BE DONE]}

{\bf Make an exclusion diagram.}

{\bf We can make predictions for pressure slopes in CO}

{\bf Coincidence of the slope changes.}

{\bf Can the height measurement help? There is still the degeneracy. Why did
Pinte et al. not have this degeneracy? How could they pin down the turbulence?}

\subsection{Including a grain size distribution}
%

{\bf [Here we will move away from a single grain size and see if we can understand
the deviations from Gaussian by a grain size distribution]}

\section{Streaming instability, clumping, and the spectral index}
\label{sec-si-clump}
%

{\bf [This section is temporarily hidden, because a lot still has to be
done on this.]}

% Whenever the dust trapping becomes so effective that the volume density of the
% dust at the midplane of the disk exceeds the gas volume density, the streaming
% instability may set in \citep{2005ApJ...620..459Y}. This produces turbulence and
% leads to clumping \citep{2007ApJ...662..627J, 2008A&A...479..883L,
%   2010ApJ...722.1437B}. If conditions are right, some of these clumps may
% gravitationally collapse and form planetesimals \citep{2007Natur.448.1022J,
%   2017ApJ...847L..12S, 2017A&A...597A..69S}. It is believed that this mechanism
% is the trigger that starts the process of the formation of planets
% \citep[see][for a review]{2014prpl.conf..547J}. If one could find observational
% evidence of this dust clumping occurring in protoplanetary disks, this would
% be a major step forward for our understanding of planet formation.
% 
% The clumps that are formed in simulations of the streaming instability are,
% however, typically substantially smaller than a pressure scale height of the
% disk \citep{2007ApJ...662..627J, 2013MNRAS.434.1460K}. This makes it very hard
% to spatially resolve these structures, even with ALMA at its highest resolution.
% 
% It might, however, be possible to find indications of particle clumping in an
% indirect way, if it leads to the formation of optically thick clouds of
% particles surrounded by optically thin regions. Even if we cannot spatially
% resolve these clouds, the spatially averaged thermal emission from such a clumpy
% medium has a different spectral slope than that of a homogeneous distribution of
% dust. If all the dust is concentrated in a multitude of small and very optically
% thick clumps, then the brightness of this unresolved clumpy medium would be the
% Planck function times the covering fraction $C$ of the clumps:
% \begin{equation}\label{eq-bright-clumpy}
% I_\nu^{\mathrm{obs}} = B_\nu(T_{\mathrm{disk}})\,C
% \end{equation}
% where the limit $C\rightarrow 0$ means concentrating all the dust in infinitely
% dense clumps so that the covering fraction is zero, and $C\rightarrow 1$ means
% having so many clumps that they multiply overlap, when seen in projection from
% above. In contrast, if all this dust is homogeneously distributed, then the
% brightness is
% \begin{equation}\label{eq-bright-homogeneous}
% I_\nu^{\mathrm{obs}} = B_\nu(T_{\mathrm{disk}})\,(1-e^{-\tau_\nu})
% \end{equation}
% with $\tau_\nu$ being the vertical optical depth of the dust in the disk.
% Reality will likely lie in between these two extreme scenarios, with
% optically thin dust interdispersed between optically thick clouds.
% 
% There are at least two ways by which one could use Eqs.~(\ref{eq-bright-clumpy},
% \ref{eq-bright-homogeneous}) to search for hints of clumpiness. The most direct
% way is to make use of the different dependencies of these two equations on
% wavelength $\lambda=c/\nu$. In the case of the clumpy medium with perfectly
% optically thick clumps (Eq.~\ref{eq-bright-clumpy}) the brightness follows the
% Planck curve, whereas for the homogeneous medium
% (Eq.~\ref{eq-bright-homogeneous}), if it is optically thin
% (i.e.~$1-e^{-\tau_\nu}\simeq \tau_\nu$), it follows the Planck curve times the
% opacity law. If we are in the Rayleigh-Jeans domain of the Planck curve, then
% these two limits are usually referred to as spectral slope
% $\alpha_{\mathrm{spec}}=2$ and $\alpha_{\mathrm{spec}}=2+\beta$, respectively,
% where $\beta$ is defined by $\kappa_\nu\propto \nu^\beta$. At ALMA band 6, for
% the low temperatures we expect in these disks at large radii, we are no longer
% strictly in the Rayleigh-Jeans domain, so the spectral slope analysis becomes
% temperature-dependent, which complicates matters a bit, but the principle
% remains the same. If the streaming instability is operating in the rings of our
% sources, and if the numbers work out (see below), then we expect to see
% $\alpha_{\mathrm{spec}}\simeq 2$ within the rings, and
% $\alpha_{\mathrm{spec}}>2$ between the rings.
% 
% There is an important caveat to this idea: the grain size and the corresponding
% opacity slope may mimick the same spectral signatures as the clumping. This is,
% in fact, the same degeneracy of the interpretation of the millimeter spectral
% slope as usual \citep[e.g.][]{2003A&A...403..323T}, just on a much smaller
% spatial scale: a scale that remains unresolved even by ALMA.  Nevertheless it is
% worthwhile to investigate if we can find clues that point to clumpiness.
% 
% In the scope of this paper we will only do a very simple analysis: we will
% verify if, under the most benign conditions, the above mentioned optically
% thick/thin clumpy medium is at all possible. This is not granted, because
% these rings are several tens of au large, and therefore it requires a lot of
% mass to make dust clumps optically thick at 1.3 millimeter wavelength.
% Furthermore, these dust clouds cannot have a density in excess of the
% Roche density, because then they would immediately gravitationally collapse.
% Therefore, we cannot make the clumps arbitrary dense and compact.
% 
% So let us construct the most optimistic model of a dust cloud. The highest
% possible density of a cloud is the Roche density, defined by
% \begin{equation}
% \rho_{\mathrm{Roche}}(r) = \frac{9}{4\pi}\frac{M_{*}}{r^3}
% \end{equation}
% where $M_{*}$ is the stellar mass and $r$ is the distance to the star. We take
% $\rho_{\mathrm{cloud}}=\rho_{\mathrm{Roche}}$. The simplest opacity model for a
% dust grain is that of \cite{1997MNRAS.291..121I}, where we omit the scattering
% part:
% \begin{equation}
% \kappa(a) = \frac{3}{4}\frac{1}{\xi\,a}\left\{\begin{matrix}
% 1 & \hbox{for} & \lambda<2\pi a \\
% 2\pi a/\lambda & \hbox{for} & \lambda>2\pi a \\
% \end{matrix}\right.
% \end{equation}
% where $a$ is the grain radius and $\xi$ is the material density of the dust
% grain ($\xi=3.6$ for amorphous olivine, $\xi=2.0$ for silica {\bf CHECK}). For a
% given wavelength $\lambda$ (in our case $\lambda=0.13\,\mathrm{cm}$ for ALMA
% band 6), the most optimistic opacity is therefore $\kappa=(3\pi/2)/\xi\lambda$,
% for all $a\le \lambda/2\pi$. For ALMA band 6 and $\xi=2\,\mathrm{g/cm}^3$ this
% is $\kappa=18.1\,\mathrm{cm}^2/\mathrm{g}$. The largest grain size to still
% have this opacity is therefore $a=\lambda/2\pi$, which for ALMA band 6 is
% $a=0.02\,\mathrm{cm}$.  We make a spherical cloud of this dust with radius
% $R$. The optical depth through the cloud center is then
% \begin{equation}
% \tau = 2R\rho_{\mathrm{cloud}}\kappa
% \end{equation}
% To obtain the effect of an optically thick/thin clumpy medium, these clouds
% have to be sufficiently optically thick, not just marginally. So let us demand
% a minimal optical depth of $\tau=4$. This gives a minimal cloud radius of
% \begin{equation}
%   R\gtrsim \frac{2}{\rho_{\mathrm{Roche}}\kappa} = 0.11\frac{1}{\rho_{\mathrm{Roche}}}
%   = 0.15\frac{r^3}{M_{*}}
% \end{equation}
% in CGS units (for ALMA band 6). Note that this scales linearly with
% $\rho_{\mathrm{cloud}}/\rho_{\mathrm{Roche}}$, so that if we choose a much less
% ``critical'' cloud density of, say, $\rho_{\mathrm{cloud}}=0.1\,\rho_{\mathrm{Roche}}$,
% then the cloud should become 10$\times$ larger and 100$\times$ more massive.
% 
% If we want these clouds to be part of an unresolved clumpy medium, they should
% be substantially smaller than the pressure scale height of the disk, and
% substantially smaller than the ALMA band 6 resolution. Let us take the example
% of the source AS 209, with $M_{*}=0.9\,M_{\odot}$, and its two high-contrast
% rings at 74 au and 120 au, respectively. This gives minimal cloud
% radii of $R\gtrsim 4\times 10^{-3}\,\mathrm{au}$ and $R\gtrsim 22\times
% 10^{-3}\,\mathrm{au}$, respectively. Clouds of size similar to these minimal
% values are indeed much smaller than the beam width, and are therefore
% unresolved.  This does not mean that there cannot be any smaller
% clumps/clouds. It only means that any smaller clouds would be optically thin,
% and we would thus not see any effect of these clumps on the unresolved
% brightness: these clouds would have a spectral slope equal to that of an
% optically thin non-clumpy model. At any rate, this estimate shows that, at least
% in principle, it is possible that the rings we see in our data consist, in fact,
% of numerous unresolved optically thick clouds interdispersed with optically thin
% dust (or without any dust in between). Given that such an unresolved, but
% optically thick, set of clumps has a lower spectral slope
% $\alpha_{\mathrm{spec}}$ than an optically thin dusty medium, an analysis of the
% measured spectral slope variations over the radial profile of the rings {\em
%   may} give us some clues as to whether clumpiness occurs or not in these
% rings, even if we cannot resolve the clumps themselves. We would expect that
% within the ring the spectral slope is closer to $\alpha_{\mathrm{spec}}\simeq 2$
% (modulo deviations from Rayleigh-Jeans) than in the inter-ring region.
% 
% However, the grain size segregation mechanism, that larger grains are more
% concentrated in the ring than smaller grains, also tends to produce this
% signature. This is particularly the case if the big grains (which are closest to
% the ring peak) are larger than $\lambda/2\pi$ while the small grains are smaller
% than $\lambda/2\pi$. To distinguish between the roles of size seggregation and
% clumping, we need an independent way to probe the grain size. Perhaps the effects
% of self-scattering induced polarization \citep{2015ApJ...809...78K} may provide
% such an avenue. 
% 
% Even if we can distinguish between size-sorting and clumping, we still have to
% clarify whether the required total dust mass for the clumping scenario is
% reasonable, and whether the required 'filling factor' is consistent with
% simulations of the streaming instability. Both questions are related. Let us go
% back to the two prominent isolated rings of AS 209. The peak linear brightness
% temperature is 3.5 $\mathrm{K}$ for the inner one and 2.7 $\mathrm{K}$ for the
% outer one. In our model the disk midplane temperature is 13 $\mathrm{K}$ for the
% inner and 10 $\mathrm{K}$ for the outer ring, which translates (for
% $\lambda=0.13\,\mathrm{cm}$) into a linear brightness temperature of
% 8 $\mathrm{K}$ and 5.6 $\mathrm{K}$, respectively. If the disk were optically
% thick, we would thus expect to observe a linear brightness temperature of
% 8 $\mathrm{K}$ and 5.6 $\mathrm{K}$, which are only about 2 times higher. 
% 
% 
% This means that the filling factor of these clouds, as projected vertically
% toward the observer, must be quite high, of the order of 50\%. {\bf [At this
%     point we can calculate the minimum mass in dust needed for this]}
% 
% In principle the disk could be warmer than we assume, leading to a larger
% saturation brightness, and thus a lower required filling factor. But given that
% a 2$\times$ higher disk temperature requires a 16$\times$ higher irradiation
% flux, we expect at most a factor of 2 uncertainty in the required filling
% factor.
% 
% Checking whether this filling factor is consistent with the expectation of
% simulations of the streaming instability is not easy. Such data are not
% published for the models describing the streaming instability so far.  We can
% only do a ``by eye'' check of the figures in those papers ({\bf give a few
%   example papers here}). Such an exercise gives us the impression that the
% typical fraction of the simulation box that is covered by the dense cluster
% regions, as projected vertically, is typically substantially less than 50\%. If
% this impression is true, then the ring emission seen in our sample cannot be
% explained by a set of spatially unresolved optically thick clouds. That does not
% imply, however, that the streaming instability does not operate in these
% rings. It only implies that if it does, it will be hidden below an optically
% thick shroud of more extended dust.
% 
% 
% 
% 
% 
% \subsection{Under which conditions should the streaming instability set in?}
% While in Section \ref{sec-si-clump} we investigated if we can recognize the
% ongoing streaming instability in our data if it is taking place in the disk, in
% this Section we will compute, with our analytic dust trapping model, under which
% condition we would {\em expect} that the streaming instability will operate in
% the first place. This depends on the total dust mass locked in the trap, the
% grain sizes and the turbulent strength. To make analytic estimates it is easiest
% to assume a single grain size, instead of a grain size distribution.
% 
% {\bf FINISH THIS SECTION}
% 
%

\section{Discussion}

\subsection{Gap depth estimates from 2-D hydrodynamic simulations}

{\bf Dong \& Fung (2017)} presented a systematic study of the gap
opening profiles for given planetary masses, and analyzed the
rings seen in scattered light of 5 sources (HD 97048, TW Hydra,
HD 169142, LkCa15 and RX J1615).

\subsection{Condition for the streaming instability}
In the literature it is often mentioned that the streaming instability requires
a dust-to-gas surface density ratio of
$\Sigma_{\mathrm{d}}/\Sigma_{\mathrm{g}}\gtrsim 0.02$ or higher to operate {\bf
  XXXXXXXX}. This can, however, not be directly compared to our models, because
this value of $0.02$ was found for models without any pre-determined
turbulence. The turbulence in those models was induced by the streaming
instability itself or, if the streaming instability does not operate, by the
Kelvin-Helmholtz instability ({\bf XXXX}). In the analytic model of this
Section, on the other hand, we set the turbulence strength by hand, by setting
$\alpha_{\mathrm{turb}}$ to some value. In essence, we assume that there is
another source of turbulence, such as the magnetorotational instability ({\bf
  XXXX}) or the vertical shear instability ({\bf XXXX}), that determines the
mixing of the dust in the disk.

According to \citet{2005ApJ...620..459Y} the true criterion for the onset of the
streaming instability is the ratio of dust and gas {\em volume} densities
$\rho_{\mathrm{d}}/\rho_{\mathrm{g}}\gtrsim 1$. For a given surface density
ratio $\Sigma_{\mathrm{d}}/\Sigma_{\mathrm{g}}$, the midplane volume density
ratio, for a single grain species with midplane Stokes number $\mathrm{St}\ll
1$, depends on the turbulent strength as
\begin{equation}\label{eq-dtg-sig-vs-rho}
  \frac{\rho_{\mathrm{d}}}{\rho_{\mathrm{g}}}\simeq
  \left(1+\frac{\mathrm{St}}{\alpha_{\mathrm{turb}}}\right)
  \frac{\Sigma_{\mathrm{d}}}{\Sigma_{\mathrm{g}}}
\end{equation}
The criterion of $\Sigma_{\mathrm{d}}/\Sigma_{\mathrm{g}}\gtrsim 0.02$ mentioned
in the literature thus relates to the criterion
$\rho_{\mathrm{d}}/\rho_{\mathrm{g}}\gtrsim 1$ via the turbulent strength and
the Stokes number. Given that we do not compute the turbulent strength
but prescribe it, we should rely on the more fundamental volume density
criterion of \citet{2005ApJ...620..459Y} to assess whether the dust in our
model triggers the streaming instability or not.

In our analytic setup of Section {\bf CHECK} %\ref{sec-anmodel-degree-trapping-alpha-series},
the dust ring is about 5 times narrower than the gas bump. Assuming that
initially the dust-to-gas ratio was $0.01$, this implies a dust-to-gas surface
density ratio at the peak of the ring of $0.05$, which is in excess of the usual
$0.02$ value. But as long as $\mathrm{St}\lesssim 19\alpha_{\mathrm{turb}}$,
Eq.~(\ref{eq-dtg-sig-vs-rho}) shows that the midplane volume density ratio still
stays below unity and no streaming instability sets in.

\subsection{Can a resolved ring be in fact a blend of several unresolved rings?}
Measuring the width of rings with finite-resolution observations is, of course,
limited to ring widths that are larger than the ``beam''. However, even for
cases where the ring width clearly exceeds the beam size by a factor of a few or
more, one may ask the question: is the wide ring we see truly a single wide
ring, or could it also be made up of a series of spatially unresolved rings that
are blended into a single wide ring due to the beam convolution? This question
is particularly relevant for the very wide rings of HD 163296. It is, of course,
hard to answer in general, since we have no observational means to test this.

But from the perspective of particle trapping by a pressure bump this question
can be rigorously answered. A long-lived radial pressure perturbation in a
protoplanetary disk cannot be much narrower than about a pressure scale height
$h_p(r)$. A dust ring produced by dust trapping in this pressure bump may become
rather narrow, dependent on a variety of parameters, as discussed in
Section \ref{sec-steady-state-analytic-trap-model}. But there can not be
more than a single such dust ring in each pressure bump.

For the wide rings of HD 163296 the pressure scale height, even under the most
optimistically low disk temperature (e.g.~10 K) the pressure scale height at
rings 1 and 2 are 2.4 au and 4.3 au, respectively, which correspond to fwhm
widths of 55 and 101 milliarcseconds, respectively (where we have multiplied by
2.355 to obtain the full-width at half-maximum corresponding to a gaussian with
$h_p$ as standard deviation width). Clearly the ALMA observations in band 6,
with fwhm beam size of 35 mas, spatially resolve the pressure scale height.
This means that the ring separation will be spatially resolved by ALMA,
ruling out the possibility that the wide rings are made up of a multitude of
narrow rings, at least in the dust trapping scenario.

\subsection{Dust trapping in the horseshoe region and in resonances}
The 1-D approach of the {\sf DISKLAB} models make it impossible to treat the
trapping of particles in the horseshoe region. The low-$\alpha$ models of {\bf
  Liu et al.~2018} show that this may be a significant effect. However, this
tends to produce non-axisymmetric rings: there will clearly be a dearth of dust
near the location of the planet, and a slight enhancement of dust near the L4
and L5 Lagrange points. Given that the rings we see so far in the multi-ringed
protoplanetary disks are angularly symmetric, this appears to be ruled out.
That means that the 1-D approach is justified, but we do have to ensure that we
are not in the regime where such non-axisymmetric horseshoe-trapping is
important. Generally the horseshoe region tends to gradually deplete over time,
but the time scale depends on the turbulent viscosity, with low-$\alpha$ disks
depleting slower.

Another issue is the potential for large particles to get trapped in a
resonance of a planet instead of in a classical dust trap {\bf REFERENCE;
the Dong double-gap is different, though}.
This requires high Stokes number, generally $\mathrm{St}\gg 1$ ({\bf is
  this true?}). Given that grain growth models tend to keep grains below
this limit ({\bf see e.g.~Birnstiel et al.}), this is less likely to play
a role.

\subsection{Multiple gaps with a single planet}
{\bf Dong, Li \& Li (2017)} and 
{\bf Bae, Zhu \& Hartmann (2017)} show that for very low-$\alpha$ disks
a planet can open multiple gaps in the disk rather than just a single one.
The contrast of these rings and gaps in the gas can be large (factor of 2; see
Bae et al.),
and thus clearly constitute dust traps. Dong et al.\ show that these gaps
can even be observable for very low mass planets.

{\bf [TODO: We can do a simple model of dust trapping for their case.]}

{\bf Fung \& Chiang (2017)} have shown that low mass planets in nearly inviscid
disks can avoid migration and can be responsible for disk accretion in spite
of the low viscosity.

\subsection{Interpretation of the millimeter flux in terms of grain size}
A common rule of thumb for interpreting millimeter and sub-millimeter
fluxes and intensities from thermal dust emission is that you ``observe
the grain size equal to the wavelength you observe at''. This rule is
based on the tendency of the dust opacity derived from a Mie calculation
to peak at wavelength around $\lambda\sim 2\pi a$, with $a$ the grain
radius. For $\lambda \ll 2\pi a$ the opacity becomes nearly constant
while $\lambda \gg 2\pi a$ it drops. This effect is even stronger for
the scattering opacity. The validity of this rule of thumb relies,
however, on the grain size distribution itself. For the simple
{\bf Ivezic et al.~(1997)} opacity law, it is only valid for
grain size distribution powerlaws obeying {\bf XXXXXXXXXXX}.
{\bf [TODO: Maybe here do an experiment with full opacity laws.]}

\begin{acknowledgements}
  The authors acknowledge support
  by the High Performance and Cloud Computing Group at the Zentrum f\"ur
  Datenverarbeitung of the University of T\"ubingen, the state of
  Baden-W\"urttemberg through bwHPC and the German Research Foundation (DFG)
  through grant no INST 37/935-1 FUGG. This research was initiated at the
  ``Stars, Planets and Galaxies'' meeting at the Harnackhaus in Berlin,
  April 2018. Part of this work was also funded by the DFG Forschergruppe
  FOR 2634 ``Planet Formation Witnesses and Probes: Transition Disks''.
\end{acknowledgements}

\begingroup
\bibliographystyle{aa}
\bibliography{ms}
\endgroup

\appendix

\section{Steady-state dust distribution in a ringlike trap}
\label{sec-steady-state-analytic-trap-model}
%
\subsection{Analytic approximate solution of dust trapping}
\label{sec-analytic-model-of-trapping}
%
Let us consider a narrow gas ring around the star at radius $r_0$ with a
midplane pressure given by
\begin{equation}\label{eq-gaussian-pressure-bump-repeat}
p(r) = p_0 \exp\left(-\frac{(r-r_0)^2}{2w^2}\right)
\end{equation}
where $w\ll r_0$ is the parameter setting the width of this gaussian gas ring.
We assume that the gas is turbulent with turbulent diffusion coefficient
$D$. Dust grains get trapped in this ring, and the dust will acquire a
radial density profile that is in equilibrium between the radial dust drift
pointing toward the peak of the gas pressure and radial turbulent diffusion
pointing away from that position. The radial dust drift velocity is
\citep[see e.g.][]{2010A&A...513A..79B}:
\begin{equation}\label{eq-v-radial-drift}
  v_{\mathrm{dr}} = \frac{\mathrm{St}}{1+\mathrm{St}^{2}}
  \left(\frac{d\ln p}{d\ln r}\right)\frac{c_s^2}{\Omega_Kr}
\end{equation}
where $c_s$ is the isothermal sound speed and the Stokes number $\mathrm{St}$ is
defined as
\begin{equation}\label{eq-definition-stokes-number}
\mathrm{St} = \Omega_Kt_{\mathrm{stop}}
\end{equation}
where $t_{\mathrm{stop}}$ is the stopping time of the grains. The diffusion
coefficient for the dust is \citep{2007Icar..192..588Y}:
\begin{equation}
D_{\mathrm{d}} = \frac{D}{1+\mathrm{St}^2}
\end{equation}
We take $D$ to be equal to the turbulent viscosity $\nu$ divided by the
Schmidt number $\mathrm{Sc}$, which we usually set to $\mathrm{Sc}=1$.
We use the usual $\alpha$-prescription for the turbulence:
\begin{equation}
D=\frac{\nu}{\mathrm{Sc}}= \alpha_{\mathrm{turb}}\frac{c_s^2}{\mathrm{Sc}\,\Omega_K}
\end{equation}
If $D$ is sufficiently small, the dust will get concentrated into a ring with
width $w_d$ that is substantially smaller than the width of the gas ring
$w$. In the following, we will ignore any terms
arising from the curvature of the coordinates. The steady-state radial
dift-mixing equation for the dust then becomes, in its approximate form:
\begin{equation}
  \frac{d}{dr}
  \left(\Sigma_{\mathrm{d}}v_{\mathrm{dr}}-D_{\mathrm{d}}\frac{d\Sigma_{\mathrm{d}}}{dr}\right) = 0
\end{equation}
Integrating this equation once, with integration constant zero (which
amounts to a zero net radial flux), yields
\begin{equation}\label{eq-drift-mix-equil-eq}
  \Sigma_{\mathrm{d}}v_{\mathrm{dr}} = D_{\mathrm{d}}\frac{d\Sigma_{\mathrm{d}}}{dr}
\end{equation}
From Eqs.(\ref{eq-v-radial-drift},\ref{eq-gaussian-pressure-bump-repeat}) we can express
$v_{\mathrm{dr}}$ as
\begin{equation}
v_{\mathrm{dr}} = -\left(\frac{c_s^2}{w^2\Omega_K(\mathrm{St}+\mathrm{St}^{-1})}\right)(r-r_0)
\end{equation}
With this expression we can solve Eq.~(\ref{eq-drift-mix-equil-eq}) for
$\Sigma_{\mathrm{d}}$, leading to the following simple analytic solution to the dust
trapping problem:
\begin{equation}\label{eq-analytic-sol-radial-trapping}
\Sigma_{\mathrm{d}}(r) = \Sigma_{\mathrm{d0}} \exp\left(-\frac{(r-r_0)^2}{2w_{\mathrm{d}}^2}\right)
\end{equation}
with 
\begin{equation}
  w_{\mathrm{d}} = w\, \sqrt{\frac{\Omega_KD_{\mathrm{d}}(\mathrm{St}+\mathrm{St}^{-1})}{c_s^2}}
  = w\,\sqrt{\frac{\alpha_{\mathrm{turb}}}{\mathrm{Sc}\,\mathrm{St}}}
\end{equation}
The normalization constant $\Sigma_{\mathrm{d0}}$ can be approximately expressed
in terms of the total dust mass trapped in the pressure bump:
\begin{equation}
  M_d = 2\pi \int_0^\infty \Sigma_{\mathrm{d}}(r) rdr \simeq
  2\pi r_0 \int_0^\infty \Sigma_{\mathrm{d}}(r)dr
\end{equation}
which leads to
\begin{equation}
\Sigma_{d0}\simeq \frac{M_d}{(2\pi)^{3/2}\, r_0\,w_{\mathrm{d}}}
\end{equation}
The approximation is best for narrow dust rings.

Note that this analytic solution is only valid as long as
$\alpha_{\mathrm{turb}}\ll \mathrm{Sc}\,\mathrm{St}$, or in other words as long
as $w_{\mathrm{d}}$ is substantially smaller than $w$.  This solution is, in
fact, the radial version of the vertical settling-mixing equilibrium solutions
of \citet{1995Icar..114..237D}.

Unfortunately, the condition that $\alpha_{\mathrm{turb}}\ll
\mathrm{Sc}\,\mathrm{St}$ (and equivalently $w_d\ll w$) is easily broken for
small grains and/or non-weak turbulence. This is similar to the case for
vertical settling and mixing: Small grains tend not to settle below a few
pressure scale heights. They settle down to some critical $z_{\mathrm{settle}}$,
below which they are well mixed with the gas, and above which they are strongly
depleted \citep{2004A&A...421.1075D}. \citet{2009A&A...496..597F} give an
analytic solution to the settling-mixing problem (their Eq.~19) that reproduces
the solutions of \citet{1995Icar..114..237D} for $\alpha_{\mathrm{turb}}\ll
\mathrm{Sc}\,\mathrm{St}$, but at the same time remains valid also for
$\alpha_{\mathrm{turb}}\gtrsim \mathrm{Sc}\,\mathrm{St}$, as long as
$\alpha_{\mathrm{turb}}$, $\mathrm{Sc}$ and $\mathrm{St}$ remain constant with
height above the midplane.

Given the similarity between the radial dust trapping problem and the vertical
settling problem, one can show that the radial version of the solution of
\citet{2009A&A...496..597F} reads:
\begin{equation}\label{eq-analytic-sol-radial-trapping-better}
  \Sigma_{\mathrm{d}}(r) = \Sigma_{\mathrm{d0}} \exp\Bigg[
    -\frac{\mathrm{Sc}\,\mathrm{St}}{\alpha_{\mathrm{turb}}}
     \left(\exp\left(\frac{\Delta r^2}{2w^2}\right)-1\right)
    -\frac{\Delta r^2}{2w^2}\Bigg]
\end{equation}
where we defined $\Delta r$ as
\begin{equation}
\Delta r \equiv (r-r_0)
\end{equation}
The solution Eq.~(\ref{eq-analytic-sol-radial-trapping-better}) is valid for
any value of $\alpha_{\mathrm{turb}}/\mathrm{Sc}\,\mathrm{St}$, as long as
$\alpha_{\mathrm{turb}}$, $\mathrm{Sc}$ and $\mathrm{St}$ remain constant along
the radial width of the dust trap, and as long as $w\ll r_0$, to prevent
geometric terms from the cilindrical coordinates from dominating.  One can
easily verify that Eq.~(\ref{eq-analytic-sol-radial-trapping-better}) reproduces
the simpler Gaussian solution Eq.~(\ref{eq-analytic-sol-radial-trapping}) for
$\Delta r\ll w$. One can also verify that for $\alpha_{\mathrm{turb}}\gg
\mathrm{Sc}\,\mathrm{St}$ the shape of $\Sigma_{\mathrm{d}}(r)$ follows the
shape of the gas pressure profile $p(r)$ (Eq.~\ref{eq-gaussian-pressure-bump-repeat}).

The solution Eq.~(\ref{eq-analytic-sol-radial-trapping-better}), like its
simplified version Eq.~(\ref{eq-analytic-sol-radial-trapping}), assumes that the
dust is always near the midplane of the disk, so that one can ignore the
vertical extent, and thereby ignore any variations in $\mathrm{St}$. In other
words, we assume that the vertical scale height of the dust layer $h_d$ is much
less than the gas pressure scale height $h_p$. As it turns out, this condition
$h_d\ll h_p$ is satisfied if $\alpha_{\mathrm{turb}}\ll
\mathrm{Sc}\,\mathrm{St}$, which is the same condition that leads to $w_d\ll w$
(at least for $\mathrm{St}\ll 1$, which is what we assume).
In fact, the radial dust ring width compared to the gas ring width is the
same as the vertical dust layer thickness compared to the gas disk thickness:
\begin{equation}
\frac{w_d}{w} = \frac{h_d}{h_p}
\end{equation}
In other words: the dust behaves radially the same way as vertically. The
vertical dust settling can be regarded as dust trapping in vertical direction,
or the radial dust trapping can be seen as radial settling of dust toward
the pressure peak.

This also means that whenever the simplified Gaussian analytic solution
(Eq.~\ref{eq-analytic-sol-radial-trapping}) ceases to be valid, so does the
simplified vertical Gaussian solution of \citet{1995Icar..114..237D}. Or in
other words, whenever the more sophisticated analytic solution
Eq.~(\ref{eq-analytic-sol-radial-trapping-better}) should be used instead of the
simplified version, the assumption that the dust is settled near the midplane
ceases to hold. Strictly speaking this invalidates
Eq.~(\ref{eq-analytic-sol-radial-trapping-better}), since this was derived under
the assumption that the dust is all located near the midplane. But given that
most of the dust mass is still close to the midplane, even for
$\alpha_{\mathrm{turb}}\gtrsim \mathrm{Sc}\,\mathrm{St}$, we are confident that
Eq.~(\ref{eq-analytic-sol-radial-trapping-better}) still remains approximately
valid.

\begin{figure}
\centerline{\includegraphics[width=0.47\textwidth]{analytic_trap_example.pdf}}
\caption{\label{fig-analytic-model-example}The analytic solution shown for three
  different grain sizes: small ($0.1\,\mu\mathrm{m}$), middle middle
  ($10\,\mu\mathrm{m}$) and large ($1000\,\mu\mathrm{m}$). The normalizations of
  the three curves are arbitrary, optimized for clarity. The solid (blue) curves
  show the full analytic solution
  (Eq.~\ref{eq-analytic-sol-radial-trapping-better}).  The dotted (green) curves
  show the simplified analytic solution
  (Eq.~\ref{eq-analytic-sol-radial-trapping}). The dashed (red) curves show the
  dust profile {\em if} the dust-to-gas ratio would be constant, which shows the
  shape of the gas pressure bump. The absense of the green dotted curve for the
  case of $1000\,\mu\mathrm{m}$ grains (the narrowest peak) is because for these
  large grains the simple analytic solution nearly perfectly overlaps with the
  full analytic solution. The parameters are chosen for AS 209, ring 1,
  with $M_{*}=0.912\,M_{\odot}$, $T_{\mathrm{gas}}=12.9\,\mathrm{K}$. The gas
  surface density is taken to be $5\,\mathrm{g}/\mathrm{cm}^2$ at the position
  of the ring, the turbulence parameter was set to
  $\alpha_{\mathrm{turb}}=2\times 10^{-4}$, the dust material density was chosen
  to be $2\,\mathrm{g}/\mathrm{cm}^3$ consistent with slightly porous silicate
  dust aggregates. The standard-deviation width of the gas pressure bump was set
  to $w=9.7\,\mathrm{au}$ which equals two times the gas pressure scale height.}
\end{figure}

The typical shapes of the analytic dust surface density profiles is shown in
Fig.~\ref{fig-analytic-model-example} for three different grain sizes from small
to large. For the largest grains, which are the most strongly trapped near the
pressure peak, and hence have the narrowest peak in
Fig.~\ref{fig-analytic-model-example}, the full analytic solution
(Eq.~\ref{eq-analytic-sol-radial-trapping-better}) nearly perfectly overlaps
with the simplified (Gaussian) analytic solution
(Eq.~\ref{eq-analytic-sol-radial-trapping}). For large grains the simplified
analytic solution is therefore sufficiently accurate. But in this parameter
regime the dust rings are much narrower than the gas pressure bump ($w_d\ll w$).

For the tiny grains, which are well coupled to the gas, and hence have the
widest peak in Fig.~\ref{fig-analytic-model-example}, the dust trapping is very
inefficient. Between 60 and 90 au the dust surface density follows nearly
perfectly the shape of the gas bump, i.e.\ keeping the dust-to-gas ratio nearly
constant. Only for $r\lesssim 60\,\mathrm{au}$ and $r\gtrsim 90\,\mathrm{au}$
the full solution drops below a constant dust-to-gas ratio. To good approximation
one can thus say that a dust ring made up of these tiny grains has a width
that equals the width of the gas pressure bump. Note that in this regime the
simplified analytic model (Eq.~\ref{eq-analytic-sol-radial-trapping}) fails
miserably, as can be seen by the nearly horizontal green dotted curve in
Fig.~\ref{fig-analytic-model-example}.

The intermediate size grains, corresponding to the intermediate curve (marked
with $100\,\mu\mathrm{m}$) in Fig.~\ref{fig-analytic-model-example}, have a
behavior between the above two extremes. The simplified analytic solution is a
bit wider than the more accurate solution. The shape of the more accurate
solution deviates from Gaussian. However, the deviation from Gaussian shape of
the model is opposite to the deviation from Gaussian shape in the observations
(Fig.~\ref{fig-obs-gaussfits}). The observed data exceed the best-fit Gaussian
in the flanks, while the model drops below the best-fit Gaussian in the flanks.

Whether this deviation from Gaussian is detectable in the ALMA data is,
however, not so clear. For intermediate size grains the non-Gaussian profile
can be easily mistaken for a Gaussian profile with a narrower width than
both the gas bump and the simple analytic solution.

\subsection{Resulting profiles for a size distribution}
\label{sec-anmodel-profiles}
%

{\bf [THIS SECTION NEEDS HEAVY EDITING]}

The analytic model of the dust trapping for a single grain size speaks for
itself. But how does it look for a grain size distribution? Since the trapping
width $w_{\mathrm{d}}$ for the different grain sizes is different, we can only
impose the grain size distribution for the total (radially integrated) dust
mass. We define the size distribution according to the following powerlaw
\begin{equation}
m(a)\frac{dN}{d\ln a} = \frac{dM}{d\ln a} \propto a^{q}
\end{equation}
where $a$ is the grain size, $m(a)$ the corresponding grain mass, $N$ the
cumulative particle number and $M$ the cumulative dust mass. The parameter $q$
is the size distribution powerlaw coefficient, and it is $q=1/2$ for the usual
MRN distribution (this corresponds to $dN/da\propto a^{-3.5}$). We also need to
define limits $a_{\mathrm{min}}$ and $a_{\mathrm{max}}$. The size distribution
is then normalized such that its integral over $d\ln(a)$ is the total dust mass
$M$. The radial surface density solution,
Eq.~(\ref{eq-analytic-sol-radial-trapping}), then becomes:
\begin{equation}\label{eq-analytic-sol-radial-trapping-sizedistr}
\frac{d\Sigma_{\mathrm{d}}(r)}{d\ln a} = \frac{1}{(2\pi)^{3/2}r_0
    w_{\mathrm{d}}(a)}\frac{dM}{d\ln a}
\exp\left(-\frac{(r-r_0)^2}{2w_{\mathrm{d}}(a)^2}\right)
\end{equation}

In Fig.~\ref{fig-anmodel-distribution-sigma} we show the radial profiles for an
MRN size distribution between $a_{\mathrm{min}}=10^{-2}\,\mathrm{cm}$ and
$a_{\mathrm{min}}=1\,\mathrm{cm}$ and a total dust mass of $M=1.3\times
10^{-5}\,M_\odot$. The dust material density is taken to be
$\xi_{\mathrm{dust}}=2\,\mathrm{g}/\mathrm{cm}^3$. We sample this size
distribution with 10 grain sizes, evenly distributed in $\ln(a)$. These are the
10 curves in Fig.~\ref{fig-anmodel-distribution-sigma}. These surface densities
are defined as being the local grain size distribution integrated over the width
of the grain size bin:
$\Sigma_{\mathrm{bin}}=\int_{\mathrm{bin}}(d\Sigma_{\mathrm{dust}}/da)da$. The
stellar mass is $M_{*}=0.9\,M_\odot$ and we set $\mathrm{Sc}=1$.

{\bf MUST ALSO SPECIFY GAS DENSITY, TO GET STOKES NUMBERS. MUST SAY
(AND CHECK) THAT EPSTEIN REGIME.}

\begin{figure}
\centerline{\includegraphics[width=0.5\textwidth]{analytic_trapping_sigma_profiles.pdf}}
\caption{\label{fig-anmodel-distribution-sigma}Results of the analytic dust
  trapping model for an MRN powerlaw grain size distribution between
  $a=0.01\,\mathrm{cm}$ and $a=1\,\mathrm{cm}$. The gas pressure bump has a FWHM
  of 70 au and a peak at $r=77\,\mathrm{au}$. Shown here are the surface density
  profiles for 10 logarithmically binned grain sizes. The curve with the
  narrowest peak corresponds to a grain size of $a=1\,\mathrm{cm}$, while that
  with the widest peak (and lowest maximum value) corresponds to a grain size of
  $a=0.01\,\mathrm{cm}$. The intermediate curves correspond to $a=0.0167$,
  0.0278, 0.0464, 0.0774, 0.1291, 0.2154, 0.3593 and 0.5995 $\mathrm{cm}$
  respectively. The dashed curve shows the total surface density of the dust,
  which is, by definition of the bin-integrated surface density, the sum of all
  10 curves. For a description of the model setup and model parameters, see
  Section \ref{sec-anmodel-profiles}.}
\end{figure}

As expected, the largest grains are concentrated the most and the smallest ones
the least. The largest grains, however, have the largest peak value of their
surface density at the exact location of the gas pressure peak.

\begin{figure}
\centerline{\includegraphics[width=0.5\textwidth]{analytic_trapping_tau_profiles.pdf}}
\caption{\label{fig-anmodel-distribution-tau}As Fig.~\ref{fig-anmodel-distribution-sigma},
  but now the vertical optical depth at $\lambda=1.3\,\mathrm{mm}$ (ALMA band 6)
  is shown. The opacity model of \citet{1997MNRAS.291..121I} was used to compute these
from the size distribution.} 
\end{figure}

When it comes to the contribution to the optical depth, however, the situation
looks a bit different. Let us assume a very simple dust opacity law, which
roughly mimicks the real behavior of dust opacities: the model of
\citet{1997MNRAS.291..121I}, where we take the scattering opacity zero for
simplicity. In this simple but illustrative opacity model the cross
section of a grain with radius $a$ is the geometric cross section $\pi a^2$
for $2\pi a>\lambda$ (with $\lambda$ the wavelength), while it is reduced by
a factor $(2\pi a/\lambda)$ for $2\pi a<\lambda$. This means that for
$2\pi a>\lambda$ the opacity (that is, the cross section per unit mass)
is $\kappa=3/4\xi_{\mathrm{dust}} a$ where $\xi_{\mathrm{dust}}$ is the material
density, here taken to be $\xi_{\mathrm{dust}}=2\,\mathrm{g}/\cm^3$. For
$2\pi a<\lambda$ this is reduced by a factor $(2\pi a/\lambda)$ and
thus becomes independent of $a$.

It can be seen from Fig.~\ref{fig-anmodel-distribution-tau} that at the peak of
the trap all grain sizes contribute equally to the optical depth. That is
because, with our opacity model, the larger grains (the ones with $2\pi a >
\lambda$) have an opacity equal to their geometric cross section per unit dust
mass, which goes as $1/a$. The smallest grains (the ones with $a=0.01$ and
$a=0.167$) have $2\pi a < \lambda$, which means that their opacity remains
constant with $a$. Hence, for these two curves in
Fig.~\ref{fig-anmodel-distribution-tau} the contribution to the peak optical
depth drops below those of the other grains. Of course, with a more realistic
opacity model this can change somewhat, but the rough principle remains valid.

More important is to notice that the relative contributions of the grain sizes
to the peak optical depth may change if a different grain size distribution is
chosen. Here we chose the MRN size distribution ($q=1/2$), but that was a rather
arbitrary choice. Given the complexity of the dust growth processes in
protoplanetary disks, it is not at all said that the MRN distribution should be
considered a preferred one. The reader may experiment with other values of
$q$ or entirely different grain size distribution shapes altogether.

% \subsection{Effect of the degree of trapping on the appearance of the ring}
% \label{sec-anmodel-degree-trapping-alpha-series}
% %
% With this simple analytic dust trapping model we can investigate how the degree
% of trapping affects the appearance of the dust ring. For a model with a single
% grain size this is easy to predict: if the turbulence is weak enough that the
% gaussian profile of $\Sigma_{\mathrm{d}}(r)$ is narrower than the ALMA beam,
% then we observe a gaussian profile with the width of the beam. Even if the dust
% in this narrow ring would be optically thick, the observed ring would have
% sub-Planckian brightness due to beam-dilution. For stronger turbulence,
% $\Sigma_{\mathrm{d}}(r)$ is spread out wider, and we observe a gaussian profile
% with the width of the dust profile itself. As in the case of the strongly
% trapped dust, also the wide rust ring has a Gaussian profile, except that if it
% is optically thick, the observed profile is flattened at the top due to
% saturation to the Planck function.
% 
% The situation gets more complex if we have a grain size distribution. For a
% given turbulence strength $\alpha_{\mathrm{turb}}$ some grains may be
% concentrated into a ring that is narrower than the ALMA beam, while others may
% be more spread out. The strongly concentrated grain population produces a cusp
% in the brightness, which will be partly smeared out by the beam convolution,
% while the weakly concentrated grain population produces a broad underlying
% brightness bump.
% 
% To see how this looks in practice we set up a realization of our analytic
% model. The peak of the pressure is put at $r_{0}=77\,\mathrm{au}$ (corresponding
% to the inner of the two main rings of AS 209), and the width is chosen to be
% $w=30\,\mathrm{au}$ (corresponding to a FWHM of 70 au).  Given that the ALMA
% beam in band 6 has a FWHM of about 35 milliarcseconds, corresponding to 4.4 au
% for the distance to AS 209, the gas bump is much wider than the beam. The two
% main dust rings of AS 209 are about 3 beams wide, so that for our wide gas bump
% model the dust is strongly concentrated. This requires a low turbulent
% $\alpha_{\mathrm{turb}}$ or large grains. As in Section
% \ref{sec-anmodel-profiles} we choose a grain size distribution between
% $a_{\mathrm{min}}=10^{-2}\,\mathrm{cm}$ and $a_{\mathrm{min}}=1\,\mathrm{cm}$,
% following the usual MRN powerlaw of $q=1/2$ and a total dust mass of
% $M=1.3\times 10^{-5}\,M_\odot$. The dust material density is taken to be
% $\xi_{\mathrm{dust}}=2\,\mathrm{g}/\mathrm{cm}^3$. We sample this size
% distribution with 10 grain sizes, evenly distributed in $\ln(a)$. For the disk
% temperature we choose a simple powerlaw $T(r)=200\,(r/\mathrm{au})^{-1/2}$. The
% stellar mass is $M_{*}=0.9\,M_\odot$, and the source distance is taken to be
% $d=126\,\mathrm{pc}$. All these setting are taken to represent the inner of the
% two main high-contrast rings of the source AS 209. We set $\mathrm{Sc}=1$.

\begin{figure}
\centerline{\includegraphics[width=0.5\textwidth]{analytic_trapping_series_alpha.pdf}}
\caption{\label{fig-anmodel-series-alpha}Results of the analytic dust trapping
  model for various turbulent strengths, for an MRN powerlaw grain size
  distribution between $a=0.01\,\mathrm{cm}$ and $a=1\,\mathrm{cm}$. The gas
  pressure bump has a FWHM of 70 au and a peak at $r=77\,\mathrm{au}$. Shown
  here is the predicted brightness temperature at a wavelength
  $\lambda=1.3\,\mathrm{mm}$ before (dashed) and after (solid) convolution with
  the ALMA beam, assuming a source distance of $d=126\,\mathrm{pc}$. For a
  description of the model setup and model parameters, see Section
  \ref{sec-anmodel-degree-trapping-alpha-series}.}
\end{figure}

% The results for several values of $\alpha_{\mathrm{turb}}$ are shown in
% Fig.~\ref{fig-anmodel-series-alpha}. In the top panel, representing weak
% turbulence, most of the dust is trapped into a ring more narrow than the
% beam. A clear cusp can be seen in the un-convolved brightness profile,
% which is due to the strongly trapped large grains of the size distribution.
% However, the convolution washes out this cusp for the most part. This
% situation is equivalent to the simple case of a single dust species being
% trapped to a narrow unresolved ring. The only difference is that for a single
% dust size the ring is either optically thick or optically thin, while in the
% multi-size case this can be a mix. That does not make a difference for the
% observed brightness profile, which is a Gaussian either way.
% 
% For the case of the strongest turbulence, the bottom panel in
% Fig.~\ref{fig-anmodel-series-alpha}, the cusp is still seen in the un-convolved
% profile, but now the convolution does not wash it out completely. The resulting
% convolved brightness profile is clearly non-Gaussian as a result. It has a
% quasi-Gaussian cusp with broad wings. When seen with ALMA such a cuspy profile
% (narrow at the peak, broad in the wings) should be recognizable.
% 
% These findings show that if the turbulent strength is weak, the predicted dust
% ring width equals the ALMA beam width, and the brightness shape should have the
% shape of the beam profile (here assumed to be Gaussian). For stronger
% turbulence, the dust ring becomes indeed broader than the ALMA beam, but then
% the shape of the observed brightness profile may teach us something about the
% grain size distribution. Roughly speaking, if the largest grains of the size
% distribution still get trapped in a narrow ring, this shows up as a cusp in the
% brightness profile. If the smallest grains of the grain size distribution are
% too small, the predicted brightness profile would be too broad compared to the
% observed ring width. In other words: if the observed brightness profile of a
% observed ring in given source is clearly radially resolved, and if it has a
% Gaussian shape, then the most likely fit to our analytic model is that of
% a narrow size distribution (quasi a single grain size) with the right turbulent
% strength to fit the observed ring width. A broader size distribution, with the
% appropriately chosen turbulent strength, would yield a cuspy profile, not a
% Gaussian one.
% 
% The two high-contrast rings of AS 209 both appear to fit a Gaussian profile with
% 2.5 times the width of the ALMA beam (Fig.~\ref{fig-as209-rings-gauss-fit}). In
% both cases it is therefore likely, according to this analysis, that the dust
% that is trapped in these rings has a narrow size distribution instead of a broad
% MRN powerlaw distribution.

\subsection{Spectral slope variations across the ring}
\label{sec-anmodel-slope-variations}
%

{\bf [TO DO]}

\subsection{Leakage and retention time of the dust in the traps}
\label{sec-leakage}
%
In our analytic model we have assumed that the pressure bump has a perfect
Gaussian shape, meaning that the further away one goes from the peak of the
trap, the steeper the pressure gradient (and thereby the faster the dust drift)
becomes. This leads to perfect trapping: no dust can leak out of the trap.
In reality the pressure bump has a finite radial width, and smoothly
connects to the global disk structure. There will be a maximum pressure
gradient at some distance away from the peak, and beyond that the pressure
gradient declines again. If the width of the trapped dust is much narrower
than this distance, virtually no dust will get lost out of the trap. But
for small grains and/or strong turbulence, the width of the dust profile may
be large enough that at the point of largest pressure gradient there is
still a substantial amount of dust. Then the dust will leak out of the
trap.

{\bf First make estimates based on current model. Then add the gas radial
velocity by accretion.}

\subsection{Stability of gas ring}
\label{sec-gas-ring-stability}

{\bf Here we must discuss how far into the wings of the gaussian is
  stable against RWI and other instabilities. Solberg-Hoiland stability
  criterion. A Gauss may not be very stabile in its flanks. This may limit
the applicability of the full analytic solution.}

\subsection{Model degeneracies}
\label{sec-anmodel-degeneracies}
%
When trying to infer the effectiveness of dust trapping within the rings with
the analytic model of Section \ref{sec-analytic-model-of-trapping}, one notices
that there are degeneracies in the possible combinations of parameters. Most
obvious is the degeneracy between grain size and turbulence: smaller grains
and stronger turbulence may yield the same dust trapping width. But also the
width of the gas pressure bump can be compensated: a narrower bump but stronger
turbulence (or smaller grains) will yield the same ring width.

{\bf [CONTINUE THIS; SOME DEGENERACIES FOLLOW DIRECTLY FROM THE EQUATIONS.]}

\subsection{Model fits to the ALMA data}
\label{sec-anmodel-fits-to-data}
%
We now apply this model to the isolated rings in our ALMA sample (Section
\ref{sec-data}). Given the degeneracies of the model in many of the parameters,
we perform the fitting procedure in three parameters, while fixing the others
to a particular choice of values.

{\bf [CONTINUE THIS; SHOW THE FIT FIGURES; SHOW THE MCMC CORNER PLOTS]}

{\bf For HD 163296 we have constraints from the sub/superkepler data of
Teague. This limits the width of the gas bump.}

\subsubsection{Could the rings be optically thick at the center?}
\label{sec-anmodel-fits-optically-thick}
%
All rings in our sample show sub-Planckian brightness, at least for the disk
temperature assumed in our model. We have calculated the disk midplane temperature
using a simple ``flaring angle recipe''. If the disks are, however, substantially
colder (by up to a factor of 2.5 to 3) then the peak flux of the rings might be
consistent with saturated emission, i.e.\ emission expected from an optically thick
disk. So the question is: how sure are we that the rings are sub-Planckian?

To reduce the midplane temperature by a factor of $2$ the irradiation flux must
be reduced roughly by a factor of $2^4=16$. Within the flaring angle recipe this
would lead to unrealistically low flaring angle. Presently we assumed $\varphi=0.02$,
so that it would have to reduce to $\varphi=0.0013$, which is so low that we can no
longer trust the flaring angle recipe. However, maybe the outer disk lies in the
shadow of the inner disk. To see whether this can reduce the temperature to low
enough values, we perform a proper radiative transfer calculation. The results are
shown in Appendix ({\bf XXXX}). It shows that {\bf XXXXX}.

There is another issue, however. If the emission saturates to the optically thick
brightness, i.e.\ to the Planck function, then the radial profile of the ring should
be flat-topped. This is clearly not what we see in the rings in our sample, with
the possible exception of the ring of GW Lup which shows a very slight hint of that.
In general, however, it seems that the simple picture of collecting so much dust into
the ring that the ring becomes optically thick does not seem to explain the observed
radial brightness profiles. On the other hand, optical thickness can also occur
on smaller scales, as we will discuss in Section \ref{sec-si-clump}.

\subsection{Computing the dust mass trapped in the rings}
\label{sec-computing-dust-mass}
%
We can compute the dust mass trapped in each of the rings by integrating over
the measured brightness temperature profiles. However, the profiles do not go
down to zero far way from the peak, and so it is not well defined what belongs
to the ring and what not. Instead, we will compute the dust mass from the
Gaussian profile that is fitted to the ring emission. This will slightly
underestimate the dust mass, because the excess emission in the flanks of the
profiles (see Fig.~\ref{fig-obs-gaussfits}) will not be included. But there are
far more serious issues to worry about, such as our limited knowledge of the
disk temperature, the uncertainties of the dust opacities, as well as optical
depth effects. So we will estimate the mass of the best fitting Gaussian model
only.

We will assume that the intrinsic dust temperature is as given in Table
\ref{tab-gauss-params} (column $T_{\mathrm{model}}$), which was computed based on
a simple flaring angle irradiation model. We consider the Gaussian profile with
the deconvolved width, i.e.\ $w_d$ (not $\sigma$). The peak linear brightness
temperature is taken to be
\begin{equation}
A_{\mathrm{deconv}} = \frac{\sigma}{w_d}\,A
\end{equation}
where $A$, $\sigma$ and $w_d$ are from Table \ref{tab-gauss-params}. We do not
account for geometric terms in the deconvolution, which is justified as long as
$\sigma\ll r$. The deconvolved Gaussian brightness profile is then
\begin{equation}\label{eq-tbrlin-deconvolved}
T_{\mathrm{br,lin,deconv}}(r) = A_{\mathrm{deconv}}\,\exp\left(-\frac{(r-r_0)^2}{2w_d^2}\right)
\end{equation}

Next we have to worry about optical depth effects. From Table
\ref{tab-gauss-params} one can see that for most rings 
$A_{\mathrm{deconv}}$ is significantly smaller than $T_{\mathrm{model,br,lin}}$.
However, for the rings of AS 209 the ratios
$A_{\mathrm{deconv}}/T_{\mathrm{model,br,lin}}$ are around 0.5, so optical depth
effects start to play a role. Given a dust surface density profile
$\Sigma_{\mathrm{d}}(r)$ as a function of radial coordinate $r$, and
given an absorption opacity $\kappa_{\mathrm{a}}$ at the wavelength we
are measuring, the absorption optical depth $\tau_{\mathrm{a}}(r)$ is
\begin{equation}\label{eq-tau-afo-sig-kap}
\tau_{\mathrm{a}}(r) = \Sigma_{\mathrm{d}}(r)\kappa_{\mathrm{a}}
\end{equation}
We assume that scattering does not play a role. The emission from this
dusty ring is then, expressed as linear brightness temperature:
\begin{equation}
T_{\mathrm{br,lin,deconv}}(r) = \left(1-e^{-\tau_{\mathrm{a}}(r)}\right)\,T_{\mathrm{model,br,lin}}
\end{equation}
If $\tau_{\mathrm{a}}(r)\ll 1$ for all $r$ this reduces to
\begin{equation}\label{eq-int-afo-tau-bnu}
T_{\mathrm{br,lin,deconv}}(r) = \tau_{\mathrm{a}}(r)\,T_{\mathrm{model,br,lin}}
\end{equation}
But given that for some rings $A_{\mathrm{deconv}}/T_{\mathrm{model,br,lin}}
=\mathrm{max}(I_{\mathrm{obs}}(r)/B_\nu(T_{\mathrm{model}}))$ can be as large as
$\sim 0.5$, this condition is not always fulfilled.  On the other hand, these
optical depth effects cannot be too strong, because then the Gaussian shape
would flatten off at the top, which is not what we see in the observations. We
will therefore continue our computation without accounting for these (mild)
optical depth effects. This will lead to a slight underestimation of the
mass. The reader can estimate the magnitude of this underestimation by comparing
$(1-\exp(-\tau_{\mathrm{a}}))$ to $\tau_{\mathrm{a}}$.

Using Eq.~(\ref{eq-int-afo-tau-bnu}) we can, from the deconvolved
$T_{\mathrm{br,lin,deconv}}(r)$ (Eq.~\ref{eq-tbrlin-deconvolved}), solve for
$\tau_{\mathrm{a}}(r)$ and obtain
\begin{equation}
  \tau_{\mathrm{a}}(r)=\frac{A_{\mathrm{deconv}}}{T_{\mathrm{model,br,lin}}}\,
  \exp\left(-\frac{(r-r_0)^2}{2w_d^2}\right)
\end{equation}
leading with Eq.~(\ref{eq-tau-afo-sig-kap}) to
\begin{equation}
  \Sigma_{\mathrm{d}}(r)=\frac{A_{\mathrm{deconv}}}{\kappa_{\mathrm{a}}T_{\mathrm{model,br,lin}}}\,
  \exp\left(-\frac{(r-r_0)^2}{2w_d^2}\right)
\end{equation}
Integrating over $r$, ignoring geometric terms leads to
\begin{equation}\label{eq-ringmass-from-obs}
  \begin{split}
    M_{\mathrm{d}} &\simeq 2\pi r_0\int_{r_{\mathrm{in}}}^{r_{\mathrm{out}}}\Sigma_{\mathrm{d}}(r)dr\\
    &= (2\pi)^{3/2}\frac{r_0w_dA_{\mathrm{deconv}}}{\kappa_{\mathrm{a}}T_{\mathrm{model,br,lin}}}\\
    &= (2\pi)^{3/2}\frac{r_0\sigma A}{\kappa_{\mathrm{a}}T_{\mathrm{model,br,lin}}}
  \end{split}
\end{equation}

\begin{figure}
\centerline{\includegraphics[width=0.47\textwidth]{kappas_alma_band_6.pdf}}
\caption{\label{fig-opacities}The dust opacity at $\lambda=0.13\,\mathrm{cm}$
  as a function of the grain radius $a$. Plotted are the absorption opacity of
  \citet{2010A&A...512A..15R}, the corresponding scattering opacity, as well as
  the analytic absorption opacity of \citet{1997MNRAS.291..121I}.}
\end{figure}

As the final step we have to make an estimate of the absorption opacity
$\kappa_{\mathrm{a}}$. This is the biggest uncertainty because it is not well
understood what the composition of the dust is, what the grain sizes are and
what the effect of the geometry is on the opacity. We will use two estimates: on
the one hand we will employ the opacity used by \citet{2010A&A...512A..15R},
which is a mixture of astronomical silicate and amorphous carbon (see
Fig.~\ref{fig-opacities}). On the other hand we will use the dummy opacity model
of \citet{1997MNRAS.291..121I}. These two opacities serve as two extremes that
bracket the likely true value. It is known that at long wavelengths the
opacities computed with the Mie procedure from laboratory measured optical
constants are likely to be underestimates by up to a factor of 10
\citep{2017A&A...600A.123D}. The more realistic opacities of
\citep{2017A&A...600A.123D} are in between our two bracketing opacities.  For
the mass estimate we use the opacity at $\lambda=0.13\,\mathrm{cm}$ for small
grains ($a=1\mu\mathrm{m}$). For the Ricci opacity this is
$\kappa_{\mathrm{a,ri}}=1.70\,\mathrm{cm}^2/\mathrm{g}$, while for the Ivezic
opacity this is $\kappa_{\mathrm{a,ri}}=31.3\,\mathrm{cm}^2/\mathrm{g}$. The
masses estimated from Eq.~(\ref{eq-ringmass-from-obs}) using these two
bracketing opacities are listed in Table \ref{fig-obs-gaussfits}. Note that the
opacity of \citet{2010A&A...512A..15R} has a peak around a grain size of 2
$\mathrm{mm}$. That peak value happens to be very close to the Ivezic value of
$31.3\,\mathrm{cm}^2/\mathrm{g}$. We can thus also regard the Ivezic mass
estimate to be the mass if all grains have a radius of 2 mm.

\end{document}